### 2018-11-27 manuscript/manuscript.tex
version 3

\documentclass[man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Satellite-detected droughts predict plant life history},
            pdfauthor={J. Grey Monroe, Brian Gill, Kathryn Turner, \& John K. McKay},
            pdfkeywords={drought adaptation, life history evolution, remote sensing,
phylogeography, herbaria records},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Satellite-detected droughts predict plant life history}
    \author{J. Grey Monroe\textsuperscript{1,2}, Brian Gill\textsuperscript{3},
Kathryn Turner\textsuperscript{4}, \& John K. McKay\textsuperscript{2}}
    \date{}

\shorttitle{Drought and life history}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO 80523, USA\\\textsuperscript{2} College of Agriculture, Colorado State University, Fort Collins, CO 80523, USA\\\textsuperscript{3} Institute for Environment and Society, Brown University, Providence, RI 02912, USA\\\textsuperscript{4} Biology Department, Pennsylvania State University, State College, PA 16802, USA}
\keywords{drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\usepackage{lineno}

\linenumbers
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

\authornote{

Correspondence concerning this article should be addressed to J. Grey
Monroe, 307 University Ave, Fort Collins, CO 80523. E-mail:
\href{mailto:monroejg@colostate.edu}{\nolinkurl{monroejg@colostate.edu}}}

\abstract{
Explaining variation in life history strategies is a long-standing goal
of evolutionary biology. For plants, annual and perennial life histories
are thought to reflect adaptation to environments that differ in the
frequency of environmental stress such as drought. Here we test this
hypothesis in \textit{Heliophila} (Brassicaceae), a diverse genus of
flowering plants native to Africa by integrating 2192 herbaria
occurrence records with 34 years of satellite-based drought detection.
Consistent with predictions from classic life history theory, we find
that perennial \textit{Heliophila} species occur in environments where
droughts are significantly less frequent compared to annuals. These
associations are predictive while controlling for phylogeny, lending
support to the hypothesis that drought related natural selection has
influenced the distributions of these strategies. Additionally, the
difference in drought frequency between annual and perennial species
distributions is greatest during the summer and fall, which also appears
to be when annuals are in the seed phase of their life cycle based on
collection dates of annual species. Together, these finding provide
empirical support for classic hypotheses about the drivers of life
history strategy in plants - that perennials out compete annuals in
environments with less frequent drought and that annuals are adpated to
enviroments with more frequent drought by escaping drought prone seasons
as seeds. 
}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Understanding the causes and consequences of life history variation is a
longstanding goal of ecology and evolutionary biology (Cole, 1954). In
plants, life histories are especially diverse, with herbaceous species
that complete their life cycle in a number of weeks to trees that live
for thousands of years (Brown, 1996). Along this continuum an important
division exists, distinguishing annuals (i.e.~monocarpic or semelparous)
which complete their seed to seed life cycle within a single calendar
year from perennials (i.e.~polycarpic or iteroparous) which can persist
over multiple years. Annual plants flower once, set seed, senesce, and
then die, spending at least some portion of the year as a seed, where
they are relatively protected from environmental stress. In contrast,
perennial plants can continue vegetative growth after reproduction and
must survive conditions experienced during all seasons. These represent
fundamentally different life history strategies, but the ecological
factors that explain their evolution and distributions remain
empirically uresolved (Friedman \& Rubin, 2015).\\
Classical theory predict shorter life spans in environments where adult
mortality is high (Charnov \& Schaffer, 1973; Franco \& Silvertown,
1996; Stearns, 1992). In plants, this has been extended to the
hypothessis that annuality is adaptive when it allows plants to escape
drought (Schaffer \& M, 1975). Lack of water is perhaps the greatest
threat to survival during vegetatitive or reproductive growth and
annuals can remain dormant (and protected as a seed) during drought.
Thus, environments with greater seasonal drought frequency may select
for annual life histories that complete reproduction prior to drought
prone seasons. Conversely, environments with less frequent drought may
select for perennial species, which may benefit from multiple bouts of
reproduction and competitive advantage by preventing recruitment of
annual species (Corbin \& D'Antonio, 2004). These predictions have been
supported by the association of annualilty with arid environments in
wild rice (Morishima, Sano, \& Oka, 1984) and \emph{Oenothera} (Evans,
Hearn, Hahn, Spangle, \& Venable, 2005). Additionally, annual and
perennial species of \emph{Nemesia} were qualitatively associated with
winter rather and summer rainfall environments respectively (Datson,
Murray, \& Steiner, 2008) and annual species of \emph{Scorzoneroides}
were associated with environments classified as unpredictable
(Cruz-Mazo, Buide, Samuel, \& Narbona, 2009). However, whether the
history frequency of drought events indeed predicts the distributions
annual or perennial life history strategies has yet to be tested in a
phylogenetic context with suitable sample sizes of distrubution and
climate data.\\
Here we combine a long-term global dataset of satellite detected drought
events with metadata from natural collections to test these classic
hypotheses about the evolution of life history strategies within the
African endemic mustard genus, \emph{Heliophila} L. (Brassicaceae). If
annuallity is an adaptive strategy allowing plants to escape drought
prone seasons, then drought frequency should predict the distribution of
life history strategies across landscapes, and annual species should be
more commonly associated with drought prone regions than perennial
species. Furthermore, if annual species have adapted to escape drought
prone seasons as seeds, observations of annual species should be rare
during drought prone seasons. Phylogenetic relatedness can have
significant non-random effects on species distributions and life history
traits (Barrett, Harder, \& Worley, 1996), and therefore we assessed the
relationship between life history distribution and drought frequency in
a phylogenetically controlled background.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{data}{%
\subsection{Data}\label{data}}

\hypertarget{availability}{%
\subsubsection{Availability}\label{availability}}

All analyses were peformed using R. All data and the source code to
produce this manuscript are available at
\url{https://github.com/greymonroe/heliophila}.

\hypertarget{satellite-detected-drought-data}{%
\subsubsection{Satellite-detected drought
data}\label{satellite-detected-drought-data}}

Remotely sensed data is a powerful tool for characterizing seasonal
patterns in drought because it is less limited in spatial and temporal
scope and resolution than weather stations or field observations
(AghaKouchak et al., 2015). To quantify the frequency of drought during
different seasons across landscapes, we used the remotely sensed
Vegetative Health Index (VHI), which measures landscape scale reductions
in plant cover and temperature conditions characteristic of drought
(Kogan, 2001). Generated from data collected by NOAA AVHRR satellites
since 1981, the VHI combines normalized difference vegetation index
(NDVI) derived measures of vegetative stress (Vegetative Condition Index
- VCI) with temperature stress indicated by anomalies in thermal spectra
(Temperature Condition Index - TCI). The VHI of year \(y\) during week
\(w\) of \([1,52]\) at pixel \(i\) is derived from the following
equations, where \(n\) is the number of years observed.

\[VCI_{y,w,i} = 100\frac{NDVI_{y,w,i} - NDVI_{min}}{NDVI_{max} - NDVI_{min}}\]

where \(NDVI_{min} = min(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})\) and
\(NDVI_{max} = max(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})\)

\[TCI_{y,w,i} = 100\frac{T_{y,w,i} - T_{min}}{T_{max} - T_{min}}\]

where \(T_{min} = min(T_{1981,w,i}...T_{1981+n,w,i})\) and
\(T_{max} = max(T_{1981,w,i}...T_{1981+n,w,i})\)

\[VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})\] Thus, VHI
measurements are standardized according to conditions historically
observed at each locations. These measurements have been validated and
generally used for evaluating drought risk and predicting crop yields in
agriculture (e.g., Rojas, Vrieling, \& Rembold, 2011; Kogan et al.,
2016). But they also present a new tool to study seasonal patterns in
the frequency of drought across environments and to test hypotheses
about the effect of drought on ecological and evolutionary processes
(Kerr \& Ostrovsky, 2003). As such, the VHI has been applied recently to
study drought related ecology of natural species and proven useful for
predicting infraspecific variation in drought tolerance traits and genes
(Dittberner et al., 2018; Mojica et al., 2016; Monroe et al., 2018).
Here, we accessed VHI data at \(16km^2\) resolution from 1981 to 2015
(\url{https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php}) to
characterize the seasonal drought frequencies experienced by annual and
perennial \emph{Heliophila} species.

\hypertarget{life-history-data-for-heliophila}{%
\subsubsection{Life history data for
Heliophila}\label{life-history-data-for-heliophila}}

\emph{Heliophila} is a genus of flowering plants endemic to the southern
portion of Africa including the Cape Floristic and Succulent Karoo
Regions. These are among the most botanically diverse environments on
Earth and the estimated \textasciitilde{}50 \emph{Heliophila} species
are considered to be the most diverse genus of the family Brassicaceae
(Mummenhoff, Al-Shehbaz, Bakker, Linder, \& Mühlhausen, 2005). This
genus includes both perennial and annual species and this change in life
history strategy has likely arisen multiple independent times (Appel \&
Al-Shehbaz, 1997; Mummenhoff et al., 2005). Furthermore, the fine scale
climatic heterogeneity of Southern Africa is ideal for studying the
distribution of traits in relation to environmental parameters (Sayre et
al., 2013). We used life histories reported by Mummenhoff et al. (2005),
grouping species with annual or perennial life histories. Perenniality
was defined based any form of perennial life history (e.g., herbs,
shrubs, mixed, etc). We also repeated these analyses without \emph{H.
subulata} and \emph{H. minima} which have mixed life histories and found
similar results. Because the original authors didn't report the nature
of the mixed trait (Mummenhoff et al., 2005), (ie. plasticity
vs.~genetic variation), we classified these species here as perennial
since they have the capacity to maintain vegetative growth after
reproduction at least to some capacity.

\hypertarget{heliophila-occurance-records}{%
\subsubsection{Heliophila occurance
records}\label{heliophila-occurance-records}}

Botanists have collected and maintained over 350 million botanical
specimens worldwide over the past 300 years. Herbarium specimens and
their associated metadata have been used since the 1960s to study
species' geographical distributions (reviewed by Willis et al. (2017)
and Lang, Willems, Scheepens, Burbano, and Bossdorf (2018)). And as they
become digitized (Soltis, 2017), these collections have been used to
study relationships between trait distributions, geography, and climate
(Davis, Willis, Connolly, Kelly, \& Ellison, 2015; Stropp et al., 2016;
Václavı'k, Beckmann, Cord, \& Bindewald, 2017; Wolf, Zimmerman,
Anderegg, Busby, \& Christensen, 2016). To characterize the diributions
of annual and perennial \emph{Heliophila} species, all records for the
genus \emph{Heliophila} were downloaded from the Global Biodiversity
Information Facility (gbif.org) on July 21, 2018 (``GBIF occurrence
download,'' 2018).

\hypertarget{sequence-data-for-phylogeny}{%
\subsubsection{Sequence data for
phylogeny}\label{sequence-data-for-phylogeny}}

Aligned Heliophila ITS sequences were obtained from previous work by
Mandáková et al. (2012). Aethionema, Alliaria, Cardamine, Chamira, and
Rorippa ITS records from were downloaded from Genbank.

\hypertarget{analyses}{%
\subsection{Analyses}\label{analyses}}

\hypertarget{drought-frequency-calculations}{%
\subsubsection{Drought frequency
calculations}\label{drought-frequency-calculations}}

To characterize drought regimens across the distrubtions of annual and
perennial species of \emph{Heliohpila}, we calculated drought during
different seasons at the location of observations for \emph{Heliophila}
records using the VHI. Specifically, we created global maps of the
frequencies of observing drought conditions (VHI\textless{}40, NOAA)
during the winter (quarter surrounding winter solstice), spring (quarter
surrounding spring equinox), summer (quarter surrounding summer
solstice) and fall (quarter surrounding fall equinox) from 1981 to 2015.
From these maps, the drought frequency during the winter, spring,
summer, and fall were extracted for the locations of all GBIF records.

\hypertarget{filtering-of-occurance-records}{%
\subsubsection{Filtering of occurance
records}\label{filtering-of-occurance-records}}

To avoid instances with spurious location data, we filtered raw GBIF by
restricting our analyses to include only:

\begin{itemize}
\tightlist
\item
  Records for species with reported life history\\
\item
  Records with geospatial data\\
\item
  Records without known geospatial issues (i.e., coordinates reported
  are those of herbarium)\\
\item
  Records from collection sites classified as land pixels\\
\item
  Records from Africa\\
\item
  Records without duplicates (i.e., identical species, location,
  collection date)
\end{itemize}

\hypertarget{phylogeny-construction}{%
\subsubsection{Phylogeny construction}\label{phylogeny-construction}}

Out group species were aligned together with \emph{Heliophila} ITS
sequences using MAFFT. Model selection for construction of phylogeny was
performed in jModeltest2 with CIPRES. Based on this analysis,
\(GTR + L\) were selected. Ultrametric phylogeny was estimated with
branch lengths as relative time.

\hypertarget{comparison-of-drought-frequency-between-annual-and-perennial-species}{%
\subsubsection{Comparison of drought frequency between annual and
perennial
species}\label{comparison-of-drought-frequency-between-annual-and-perennial-species}}

To evaluate the hypothesis that annual and perennial life history
strategies reflect adaptations to alternative drought regimes, we tested
the corresponding prediction that the observed distributions of annual
and perennial \emph{Heliophila} species would be significantly
associated with historic drought frequency. First, we compared the
frequency of drought during the winter, spring, summer, and fall between
raw occurance records of annual and perennial species by t-tests. To
account for variation in the number of occurance records per species, we
next calculated the mean drought frequency during the winter, spring,
summer and fall for each species. The relationships between species mean
values of drought frequency during each season and life habitat (annual
or perennial) were tested using Firth's penalized-likelihood logistic
regressions. Because demographic histories caused by ancestry can
confound trait - environment associations, we then tested for the
relationships between drought frequeny and life history while
controlling for relationships between species using phylogenetic
logistic regressions.

\hypertarget{collection-dates}{%
\subsubsection{Collection dates}\label{collection-dates}}

To test the hypothesis that annual species have adapted to escape
drought prone seasons as seeds, collection dates for herbarium specimens
were compared between annual and perennial species. Comparisons of
distributions were made by Two-sample Kolmogorov-Smirnov test, t-test,
and Barlett variance test.

\hypertarget{results}{%
\section{Results}\label{results}}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/phylogeny} \caption{Species and examples of herbaria specimens of
\emph{Heliophila} (a) Phylogeny and life history strategies of species
studied. Orange circles at branch tips mark annual species and blue
circles mark perennial species. Example herbaria specimens accessed via
GBIF of (a) \emph{H. minima}, (b) \emph{H. deserticola}, (c) \emph{H.
coronopifolia} and (d) \emph{H. ephemera}. Images (a,c,d) courtesy of
The Bavarian Natural History Collections (CC BY-SA 4.0) and (b) The
London Natural History Museum (CC BY 4.0). Links to images are found in
the supplement.}\label{fig:phylogeny}
\end{figure}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/maps_boxplots} \caption{Locations of (a) annual and (b) perennial Heliophila.
Drought frequency during the (c) winter, (d) spring, (e) summer and (f)
fall. (g) Drought frequencies during each season observed at the
collection sites of \emph{Heliophila} records.}\label{fig:mapsboxplots}
\end{figure}

Out of 8670 \emph{Heliophila} GBIF recrods, 6634 were for species with
reported life history (Mummenhoff et al., 2005), 3653 had geospatial
data, 3627 did not have geospatial issues, 3460 were located on pixels
classified as land having drought measurements, 3457 were located in
Africa, 3162 were not duplicated. After all filtering steps, 2192
records for 42 species (Figure \ref{fig:phylogeny}, Table
\ref{tab:speciesmeanstable}) passed for further analyses. The number of
samples varied between species, with a mean of 52.19 samples per
species. Heliophila rigidiuscula had the most records, 201, and
Heliophila cornellsbergia the fewest, 2 (Table
\ref{tab:speciesmeanstable}).

There were clear visual differences between the distributions of the 960
annual and the 1232 perennial \emph{Heliophila} observation records (see
Figure \ref{fig:speciesmaps} for maps of individual species). While
annual species were generally found in the western regions of South
Africa and the Namibia, primarily in the Cape Floristic Region and
Succulent Karoo (Figure \ref{fig:mapsboxplots}a), the occurances of
perennials extended to the east coast of South Africa (Figure
\ref{fig:mapsboxplots}b).

The frequency of drought varied considerably across the ranges of
\emph{Heliophila} species (Figure \ref{fig:mapsboxplots}c-f). This
heterogeneity is expected, given that this is one of the most
climatically diverse regions of the Earth (Sayre et al., 2013). It is
worth noting the east to west cline in drought frequency observed during
the summer, which distinguishes the high drought frequency of the
Kalahari Sands and Namid Destert phyotogeographic regions from the low
drought frequency of the Drakensberg Mountains and Coastal Zambesian
phytogeographic regions. In the Cape phytogeographic region there was
finer scale heterogeneity in drought frequency during the summer.

Theory has predicted that annuality should be adaptive in places where
stresses such as drought are more common. Conversely, perenniality
should be adaptive in places where such stresses are less frequent. We
found that the frequency of drought was significantly higher at the
locations of occurance records for annual species. In terms of raw
observation records (Figure @\ref(fig:mapsboxplots)g), the frequency of
drought was signficantly higher at the location of annuals during the
winter (t = 10.65, p = 0.00), spring (t = 10.73, p = 0.00), summer (t =
12.67, p = 0.00), and fall (t = 15.26, p = 0.00). Because raw occurance
records do not account for variation in the number of records per
species (Table \ref{tab:speciesmeanstable})), we also tested whether
mean drought frequency values of each species were significantly
different between annuals and perennials by a Firths-penalized logistic
regression. We found that the mean drought frequencies were signficantly
higher (\(\alpha = 0.05\)) in annual species during the spring, summer,
and fall (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a). We
further tested whether annual species are found in places where droughts
occur more frequently while controlling for the phylogentic relatedness
of \emph{Heliophila} species. This is important, because environmental
differences in species distributions can be confounded with demographic
history caused by acestry. We found that while controlling for
phylogeny, the mean drought frequencies were signficantly higher
(\(\alpha = 0.05\)) in annual species during the spring, summer, and
fall (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a). These
findings indicate that common acestry alone does not explain differences
the drought frequencies experienced in the environemnts of annual and
perennial \emph{Heliophila}.

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:modelstable}Logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species the winter, spring, summer, and fall.}
\begin{tabular}{lllll}
\toprule
Predictor & \multicolumn{1}{c}{Estimate`} & \multicolumn{1}{c}{P`} & \multicolumn{1}{c}{Estimate*} & \multicolumn{1}{c}{P*}\\
\midrule
Intercept & 2.2575 & 0.1739 & 0.7231 & 0.6636\\
Winter drought freq. & -6.7484 & 0.1661 & -1.5452 & 0.7274\\ \midrule
Intercept & 4.5594 & 0.0443 & 5.0107 & 0.0534\\
Spring drought freq. & -11.7895 & 0.0423 & -12.9014 & 0.0464\\ \midrule
Intercept & 7.1742 & 0.0011 & 7.7093 & 0.0054\\
Summer drought freq. & -18.2999 & 0.0010 & -19.9056 & 0.0042\\ \midrule
Intercept & 6.4226 & 0.0029 & 7.0162 & 0.0082\\
Fall drought freq. & -19.0512 & 0.0026 & -20.8174 & 0.0067\\ \midrule
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} ` = Firth's penalized logistic regression. * = Phylogenetically constrained logistic regression. Annual species were scored as 0 and perennial species as 1.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{figure}
\centering
\includegraphics{../figures/line_and_dates.pdf}
\caption{\label{fig:lineplots}Comparison (mean +- SE) of drought frequency across
seasons measured at the GBIF records of annual and perennial species of
Heliophila. (B) Collection dates of GBIF records of annual and perennial
species of Heliophila.}
\end{figure}

The preceding results indicated that annual speices are found in
environments where droughts are signifacantly more frequent, especially
in the summer and fall. Classic life history theory hypothesizes that
annuality reflects adaptation to such environments because it allows
species to escape stressful conditions. If this is the case, we would
expect that annuals spend the drought prone seasons of summer and fall
as seeds. To test this hypothesis, we compared the dates of occurance
records between annual and perennial \emph{Heliophila} species. The
distributions reveal a considerable difference in the timing of
observation of these two life histories. In comparison to perennials,
which appear to be collected throughout the year, annuals are almost
exclusively observed during the spring (Figure \ref{fig:lineplots}b).
The differences between the distrubution of collection dates were
signicant by all tests (ks.test D = 0.25, p = 0; bartlett.test K2 =
503.18, p = 0.00) This is consistent with a model of life history in
which annual species flower in the spring, set seed, senesce, and die
before the summer. Thus, these annual species are likely to remain
dormant during the summer and fall, when drought is the strongest
predictor of the distrubutions of annual and perennial life histories
(Figure \ref{fig:lineplots}a).

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

To test the hypothesis that annual and perennial plants reflect
adaptation to alternative drought environments we examined the landscape
distribution of life history strategies in the large and diverse mustard
genus, \emph{Heliophila}. Using metadata of 2192 occurance records and a
34 year dataset of satellite-detected droughts, we tested the prediction
that annual species are more often observed in drought-prone locations
than perennial species, when controlling for phylogenetic relatedness.
We found that drought frequency is significantly different between the
distributions of annual and perennial species, with annuals being found
in environments with more frequent drought, and that this signal is
strongest during the summer. These results remain significant while
controlling for the phylogenetic relationships of \emph{Heliophila}
species, yielding support for the role that natural selection has played
in driving contemporary distributions of these alternatives strategies
in relation to drought regimes.

We cannot elimiate the possibility that counfounding traits or
environmental variables are the causative factors explaining variation
in the distributions of annual and perennial species. Nevertheless,
these results reveal quantitative support for the classic prediction
that annual species are found in environemnts that experience more
frequent drought than perennial species. These findings complement
previous reports of qualitative associations between annaulity with
enviroments characterized as having increased aridity (Evans et al.,
2005), alternative precipitation defined habitats (Datson et al., 2008;
Morishima et al., 1984), or greater unpredictability (Cruz-Mazo et al.,
2009). However, to our knowledge this is the first study to demonstrate
a signficant association between life history and drought in a
phylogenetic context informed by large scale species distribution data
and long term drought measures.

Unfortunately, herbarium collections and their associated data do not
represent systematic or random sampling of a species distribution.
Significant biases in collecting exist, which we have not necessarily
controlled for here, and may have some effect on our findings, such as a
bias toward collecting near roads or near the locations of natural
history collections (Daru et al., 2018). Future research will benefit
from systematic sampling efforts to avoid these noted biases. However,
the ecosystems of southern Africa include several biodiversity hotspots
and are among the most botanically well sampled regions on Earth (Daru
et al., 2018), suggesting that this may currently be the optimal region
for our analyses of life history distribution. Indeed, we were able to
use 2192 occurance records to study 42 species, which represents a
significant advance over relying on personal observations to
characterize species distrubtuions.

These findings support classical theoretical predictions about the
adaptive value of annual and perennial life history strategies. Taken
together, they suggest that in \emph{Heliophila}, annual species are
adapted to environments with increased summer droughts by avoiding these
seasons in a dormant seed phase of their life cycle. They also suggest
that perenniality is adaptive in environemnts where droughts are less
frequent. While most previous work has focused on describing the
evolutionary origins of annuality (Andreasen \& Baldwin, 2001; Barrett
et al., 1996; Conti, Soltis, Hardig, \& Schneider, 1999; Friedman \&
Rubin, 2015; Verboom, Linder, \& Stock, 2004) there are at least a few
other cases where perenniality appears to have arisen from an annual
acestor (Bena, Lejeune, Prosperi, \& Olivieri, 1998; Tank \& Olmstead,
2008). And while early theory predicted selection for annuality when
adult morality is high (Stearns, 1992), we also find evidence that the
transition to perenniality could be explained by historical drought
regimes. The phylogeny reveals several transitions from annual to
perennial life history (Figure \ref{fig:phylogeny}a) and that perennials
distributions extend into regions where drought frequency is low (Figure
\ref{fig:mapsboxplots}b,Figure \ref{fig:speciesmaps}). Perennials may be
able to out complete annual relatives in environments where the
infrequency of drought favors strategies that allow plants to benefit
from growth over many seasons. It may also indicate that annuals rely on
drought as a source of disturbance for seedling recruitment when
competing with perennials (Corbin \& D'Antonio, 2004). Indeed, no annual
species were observed in the low drought regions of east South Africa
\ref{fig:mapsboxplots}a,Figure \ref{fig:speciesmaps}).

These findings suggest that species with locally adapted life history
strategies could be threatened by rapidly changing drought regimes (Dai,
2011). This could have impacts on ecosystem functioning if the
composition of annual and perennial species changes as a response
(Garnier, Cordonnier, Guillerm, \& Sonié, 1997; Roumet, Urcelay, \&
Dı'az, 2006). Furthermore, the frequency of drought may be an important
factor when considering the use of perennial cropping sysyems (Lelièvre
\& Volaire, 2009; Parry, Flexas, \& Medrano, 2005).

In conclusion, we find strong support for classic life history theory
which predicts that annuality is adaptive environments where droughts
occur more frequently. Additionally, we report evidence consistent with
a life history model in annuals in which they escape drought prone
seasons during the seed phase of their life cycle. Finally, we find
evidence that the distributions of perennial lineages may indicate a
competitive advantage in areas where droughts are infrequent. More
broadly, this work demonstrates the power of emerging data to address
outstanding classic hypotheses in ecology and evolution.

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

We thank \_\_ for generous feedback that improved the quality of this
work. This research was supported by NSF Award 1701918 and USDA-NIFA
Award 2014-38420-21801 to JGM.

\hypertarget{references}{%
\section{References}\label{references}}

\newpage

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-aghakouchak2015remote}{}%
AghaKouchak, A., Farahmand, A., Melton, F., Teixeira, J., Anderson, M.,
Wardlow, B. D., \& Hain, C. (2015). Remote sensing of drought: Progress,
challenges and opportunities. \emph{Reviews of Geophysics},
\emph{53}(2), 452--480.

\leavevmode\hypertarget{ref-R-geiger_a}{}%
Alfaro, M., Santini, F., Brock, C., Alamillo, H., Dornburg, A., Rabosky,
D., \ldots{} Harmon, L. (2009). Nine exceptional radiations plus high
turnover explain species diversity in jawed vertebrates.
\emph{Proceedings of the National Academy of Sciences of the United
States of America}, \emph{106}, 13410--13414.

\leavevmode\hypertarget{ref-andreasen2001unequal}{}%
Andreasen, K., \& Baldwin, B. G. (2001). Unequal evolutionary rates
between annual and perennial lineages of checker mallows (sidalcea,
malvaceae): Evidence from 18S--26S rDNA internal and external
transcribed spacers. \emph{Molecular Biology and Evolution},
\emph{18}(6), 936--944.

\leavevmode\hypertarget{ref-appel1997generic}{}%
Appel, O., \& Al-Shehbaz, I. A. (1997). Generic limits and taxonomy of
hornungia, pritzelago, and hymenolobus (brassicaceae). \emph{Novon},
338--340.

\leavevmode\hypertarget{ref-R-papaja}{}%
Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with
R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}

\leavevmode\hypertarget{ref-barrett1996comparative}{}%
Barrett, S. C. H., Harder, L. D., \& Worley, A. C. (1996). The
comparative biology of pollination and mating in flowering plants.
\emph{Phil. Trans. R. Soc. Lond. B}, \emph{351}(1345), 1271--1280.

\leavevmode\hypertarget{ref-R-Matrix}{}%
Bates, D., \& Maechler, M. (2018). \emph{Matrix: Sparse and dense matrix
classes and methods}. Retrieved from
\url{https://CRAN.R-project.org/package=Matrix}

\leavevmode\hypertarget{ref-bena1998molecular}{}%
Bena, G., Lejeune, B., Prosperi, J.-M., \& Olivieri, I. (1998).
Molecular phylogenetic approach for studying life-history evolution: The
ambiguous example of the genus medicago l. \emph{Proceedings of the
Royal Society of London B: Biological Sciences}, \emph{265}(1401),
1141--1151.

\leavevmode\hypertarget{ref-brown1996oldlist}{}%
Brown, P. M. (1996). OLDLIST: A database of maximum tree ages.
\emph{Tree Rings, Environment, and Humanity. Radiocarbon}, \emph{1996},
727--731.

\leavevmode\hypertarget{ref-charnov1973life}{}%
Charnov, E. L., \& Schaffer, W. M. (1973). Life-history consequences of
natural selection: Cole's result revisited. \emph{The American
Naturalist}, \emph{107}(958), 791--793.

\leavevmode\hypertarget{ref-cole1954population}{}%
Cole, L. C. (1954). The population consequences of life history
phenomena. \emph{The Quarterly Review of Biology}, \emph{29}(2),
103--137.

\leavevmode\hypertarget{ref-conti1999phylogenetic}{}%
Conti, E., Soltis, D. E., Hardig, T. M., \& Schneider, J. (1999).
Phylogenetic relationships of the silver saxifrages (saxifraga, sect.
Ligulatae haworth): Implications for the evolution of substrate
specificity, life histories, and biogeography. \emph{Molecular
Phylogenetics and Evolution}, \emph{13}(3), 536--555.

\leavevmode\hypertarget{ref-corbin2004competition}{}%
Corbin, J. D., \& D'Antonio, C. M. (2004). Competition between native
perennial and exotic annual grasses: Implications for an historical
invasion. \emph{Ecology}, \emph{85}(5), 1273--1283.

\leavevmode\hypertarget{ref-cruz2009molecular}{}%
Cruz-Mazo, G., Buide, M., Samuel, R., \& Narbona, E. (2009). Molecular
phylogeny of scorzoneroides (asteraceae): Evolution of heterocarpy and
annual habit in unpredictable environments. \emph{Molecular
Phylogenetics and Evolution}, \emph{53}(3), 835--847.

\leavevmode\hypertarget{ref-dai2011drought}{}%
Dai, A. (2011). Drought under global warming: A review. \emph{Wiley
Interdisciplinary Reviews: Climate Change}, \emph{2}(1), 45--65.

\leavevmode\hypertarget{ref-daru2018widespread}{}%
Daru, B. H., Park, D. S., Primack, R. B., Willis, C. G., Barrington, D.
S., Whitfeld, T. J., \ldots{} others. (2018). Widespread sampling biases
in herbaria revealed from large-scale digitization. \emph{New
Phytologist}, \emph{217}(2), 939--955.

\leavevmode\hypertarget{ref-datson2008climate}{}%
Datson, P., Murray, B., \& Steiner, K. (2008). Climate and the evolution
of annual/perennial life-histories in nemesia (scrophulariaceae).
\emph{Plant Systematics and Evolution}, \emph{270}(1-2), 39--57.

\leavevmode\hypertarget{ref-davis2015herbarium}{}%
Davis, C. C., Willis, C. G., Connolly, B., Kelly, C., \& Ellison, A. M.
(2015). Herbarium records are reliable sources of phenological change
driven by climate and provide novel insights into species' phenological
cueing mechanisms. \emph{American Journal of Botany}, \emph{102}(10),
1599--1609.

\leavevmode\hypertarget{ref-dittberner2018natural}{}%
Dittberner, H., Korte, A., Mettler-Altmann, T., Weber, A., Monroe, G.,
\& Meaux, J. de. (2018). Natural variation in stomata size contributes
to the local adaptation of water-use efficiency in arabidopsis thaliana.
\emph{bioRxiv}, 253021.

\leavevmode\hypertarget{ref-R-geiger_b}{}%
Eastman, J., Alfaro, M., Joyce, P., Hipp, A., \& Harmon, L. (2011). A
novel comparative method for identifying shifts in the rate of character
evolution on trees. \emph{Evolution}, \emph{65}, 3578--3589.

\leavevmode\hypertarget{ref-evans2005climate}{}%
Evans, M. E., Hearn, D. J., Hahn, W. J., Spangle, J. M., \& Venable, D.
L. (2005). Climate and life-history evolution in evening primroses
(oenothera, onagraceae): A phylogenetic comparative analysis.
\emph{Evolution}, \emph{59}(9), 1914--1927.

\leavevmode\hypertarget{ref-franco1996life}{}%
Franco, M., \& Silvertown, J. (1996). Life history variation in plants:
An exploration of the fast-slow continuum hypothesis. \emph{Phil. Trans.
R. Soc. Lond. B}, \emph{351}(1345), 1341--1348.

\leavevmode\hypertarget{ref-friedman2015all}{}%
Friedman, J., \& Rubin, M. J. (2015). All in good time: Understanding
annual and perennial strategies in plants. \emph{American Journal of
Botany}, \emph{102}(4), 497--499.

\leavevmode\hypertarget{ref-garnier1997specific}{}%
Garnier, E., Cordonnier, P., Guillerm, J.-L., \& Sonié, L. (1997).
Specific leaf area and leaf nitrogen concentration in annual and
perennial grass species growing in mediterranean old-fields.
\emph{Oecologia}, \emph{111}(4), 490--498.

\leavevmode\hypertarget{ref-gbifdownload}{}%
GBIF occurrence download. (2018, July). Retrieved from
\url{https://doi.org/10.15468/dl.xmwqxm}

\leavevmode\hypertarget{ref-R-mvtnorm}{}%
Genz, A., \& Bretz, F. (2009). \emph{Computation of multivariate normal
and t probabilities}. Heidelberg: Springer-Verlag.

\leavevmode\hypertarget{ref-R-MCMCglmm}{}%
Hadfield, J. D. (2010). MCMC methods for multi-response generalized
linear mixed models: The MCMCglmm R package. \emph{Journal of
Statistical Software}, \emph{33}(2), 1--22. Retrieved from
\url{http://www.jstatsoft.org/v33/i02/}

\leavevmode\hypertarget{ref-R-geiger_d}{}%
Harmon, L., Weir, J., Brock, C., Glor, R., \& Challenger, W. (2008).
GEIGER: Investigating evolutionary radiations. \emph{Bioinformatics},
\emph{24}, 129--131.

\leavevmode\hypertarget{ref-R-logistf}{}%
Heinze, G., \& Ploner, M. (2018). \emph{Logistf: Firth's bias-reduced
logistic regression}. Retrieved from
\url{https://CRAN.R-project.org/package=logistf}

\leavevmode\hypertarget{ref-R-purrr}{}%
Henry, L., \& Wickham, H. (2018). \emph{Purrr: Functional programming
tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}

\leavevmode\hypertarget{ref-R-raster}{}%
Hijmans, R. J. (2018). \emph{Raster: Geographic data analysis and
modeling}. Retrieved from
\url{https://CRAN.R-project.org/package=raster}

\leavevmode\hypertarget{ref-R-phylolm}{}%
Ho, L. S. T., \& Ane, C. (2014). A linear-time algorithm for gaussian
and non-gaussian trait evolution models. \emph{Systematic Biology},
\emph{63}, 397--408.

\leavevmode\hypertarget{ref-kerr2003space}{}%
Kerr, J. T., \& Ostrovsky, M. (2003). From space to species: Ecological
applications for remote sensing. \emph{Trends in Ecology \& Evolution},
\emph{18}(6), 299--305.

\leavevmode\hypertarget{ref-kogan2016modelling}{}%
Kogan, F., Guo, W., Strashnaia, A., Kleshenko, A., Chub, O., \&
Virchenko, O. (2016). Modelling and prediction of crop losses from noaa
polar-orbiting operational satellites. \emph{Geomatics, Natural Hazards
and Risk}, \emph{7}(3), 886--900.

\leavevmode\hypertarget{ref-kogan2001operational}{}%
Kogan, F. N. (2001). Operational space technology for global vegetation
assessment. \emph{Bulletin of the American Meteorological Society},
\emph{82}(9), 1949--1964.

\leavevmode\hypertarget{ref-lang2018using}{}%
Lang, P. L., Willems, F. M., Scheepens, J., Burbano, H. A., \& Bossdorf,
O. (2018). \emph{Using herbaria to study global environmental change}.
PeerJ Preprints.

\leavevmode\hypertarget{ref-lelievre2009current}{}%
Lelièvre, F., \& Volaire, F. (2009). Current and potential development
of perennial grasses in rainfed mediterranean farming systems.
\emph{Crop Science}, \emph{49}(6), 2371--2378.

\leavevmode\hypertarget{ref-mandakova2012whole}{}%
Mandáková, T., Mummenhoff, K., Al-Shehbaz, I. A., Mucina, L.,
Mühlhausen, A., \& Lysak, M. A. (2012). Whole-genome triplication and
species radiation in the southern african tribe heliophileae
(brassicaceae). \emph{Taxon}, \emph{61}(5), 989--1000.

\leavevmode\hypertarget{ref-Mojica2016}{}%
Mojica, J. P., Mullen, J., Lovell, J. T., Monroe, J. G., Paul, J. R.,
Oakley, C. G., \& McKay, J. K. (2016). Genetics of water use physiology
in locally adapted Arabidopsis thaliana. \emph{Plant Science}.
doi:\href{https://doi.org/10.1016/j.plantsci.2016.03.015}{10.1016/j.plantsci.2016.03.015}

\leavevmode\hypertarget{ref-monroe2018drought}{}%
Monroe, J. G., Powell, T., Price, N., Mullen, J., Howard, A., Evans, K.,
\ldots{} McKay, J. (2018). Drought adaptation in nature by extensive
genetic loss-of-function. \emph{bioRxiv}, 372854.

\leavevmode\hypertarget{ref-morishima1984differentiation}{}%
Morishima, H., Sano, Y., \& Oka, H. (1984). Differentiation of perennial
and annual types due to habitat conditions in the wild riceOryza
perennis. \emph{Plant Systematics and Evolution}, \emph{144}(2),
119--135.

\leavevmode\hypertarget{ref-mummenhoff2005phylogeny}{}%
Mummenhoff, K., Al-Shehbaz, I. A., Bakker, F. T., Linder, H. P., \&
Mühlhausen, A. (2005). Phylogeny, morphological evolution, and
speciation of endemic brassicaceae genera in the cape flora of southern
africa. \emph{Annals of the Missouri Botanical Garden}, 400--424.

\leavevmode\hypertarget{ref-R-bindrcpp}{}%
Müller, K. (2018). \emph{Bindrcpp: An 'rcpp' interface to active
bindings}. Retrieved from
\url{https://CRAN.R-project.org/package=bindrcpp}

\leavevmode\hypertarget{ref-R-tibble}{}%
Müller, K., \& Wickham, H. (2018). \emph{Tibble: Simple data frames}.
Retrieved from \url{https://CRAN.R-project.org/package=tibble}

\leavevmode\hypertarget{ref-R-caper}{}%
Orme, D., Freckleton, R., Thomas, G., Petzoldt, T., Fritz, S., Isaac,
N., \& Pearse, W. (2018). \emph{Caper: Comparative analyses of
phylogenetics and evolution in r}. Retrieved from
\url{https://CRAN.R-project.org/package=caper}

\leavevmode\hypertarget{ref-R-ape}{}%
Paradis, E., \& Schliep, K. (2018). Ape 5.0: An environment for modern
phylogenetics and evolutionary analyses in R. \emph{Bioinformatics},
\emph{xx}, xxx--xxx.

\leavevmode\hypertarget{ref-parry2005prospects}{}%
Parry, M., Flexas, J., \& Medrano, H. (2005). Prospects for crop
production under drought: Research priorities and future directions.
\emph{Annals of Applied Biology}, \emph{147}(3), 211--226.

\leavevmode\hypertarget{ref-R-sp}{}%
Pebesma, E. J., \& Bivand, R. S. (2005). Classes and methods for spatial
data in R. \emph{R News}, \emph{5}(2), 9--13. Retrieved from
\url{https://CRAN.R-project.org/doc/Rnews/}

\leavevmode\hypertarget{ref-R-coda}{}%
Plummer, M., Best, N., Cowles, K., \& Vines, K. (2006). CODA:
Convergence diagnosis and output analysis for mcmc. \emph{R News},
\emph{6}(1), 7--11. Retrieved from
\url{https://journal.r-project.org/archive/}

\leavevmode\hypertarget{ref-R-gee}{}%
R by Thomas Lumley, V. J. C. P. to, \& author., B. R. N. that
maintainers are not available to give advice on using a package they did
not. (2015). \emph{Gee: Generalized estimation equation solver}.
Retrieved from \url{https://CRAN.R-project.org/package=gee}

\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. (2018). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.
Retrieved from \url{https://www.R-project.org/}

\leavevmode\hypertarget{ref-R-phytools}{}%
Revell, L. J. (2012). Phytools: An r package for phylogenetic
comparative biology (and other things). \emph{Methods in Ecology and
Evolution}, \emph{3}, 217--223.

\leavevmode\hypertarget{ref-R-maps}{}%
Richard A. Becker, O. S. code by, Ray Brownrigg. Enhancements by Thomas
P Minka, A. R. W. R. version by, \& Deckmyn., A. (2018). \emph{Maps:
Draw geographical maps}. Retrieved from
\url{https://CRAN.R-project.org/package=maps}

\leavevmode\hypertarget{ref-rojas2011assessing}{}%
Rojas, O., Vrieling, A., \& Rembold, F. (2011). Assessing drought
probability for agricultural areas in africa with coarse resolution
remote sensing imagery. \emph{Remote Sensing of Environment},
\emph{115}(2), 343--352.

\leavevmode\hypertarget{ref-roumet2006suites}{}%
Roumet, C., Urcelay, C., \& Dı'az, S. (2006). Suites of root traits
differ between annual and perennial species growing in the field.
\emph{New Phytologist}, \emph{170}(2), 357--368.

\leavevmode\hypertarget{ref-sayre2013new}{}%
Sayre, R. G., Comer, P., Hak, J., Josse, C., Bow, J., Warner, H.,
\ldots{} others. (2013). A new map of standardized terrestrial
ecosystems of africa. \emph{African Geographical Review}.

\leavevmode\hypertarget{ref-schaffer1975selection}{}%
Schaffer, W., \& M, G. (1975). Selection for optimal life histories in
plants. \emph{Ecology and Evolution of Communities.}, 142--157.

\leavevmode\hypertarget{ref-R-geiger_c}{}%
Slater, G., Harmon, L., Wegmann, D., Joyce, P., Revell, L., \& Alfaro,
M. (2012). Fitting models of continuous trait evolution to incompletely
sampled comparative data using approximate bayesian computation.
\emph{Evolution}, \emph{66}, 752--762.

\leavevmode\hypertarget{ref-R-diagram}{}%
Soetaert, K. (2017). \emph{Diagram: Functions for visualising simple
graphs (networks), plotting flow diagrams}. Retrieved from
\url{https://CRAN.R-project.org/package=diagram}

\leavevmode\hypertarget{ref-R-shape}{}%
Soetaert, K. (2018). \emph{Shape: Functions for plotting graphical
shapes, colors}. Retrieved from
\url{https://CRAN.R-project.org/package=shape}

\leavevmode\hypertarget{ref-soltis2017digitization}{}%
Soltis, P. S. (2017). Digitization of herbaria enables novel research.
\emph{American Journal of Botany}, \emph{104}(9), 1281--1284.

\leavevmode\hypertarget{ref-stearns1992evolution}{}%
Stearns, S. C. (1992). \emph{The evolution of life histories}.

\leavevmode\hypertarget{ref-stropp2016mapping}{}%
Stropp, J., Ladle, R. J., M. Malhado, A. C., Hortal, J., Gaffuri, J., H.
Temperley, W., \ldots{} Mayaux, P. (2016). Mapping ignorance: 300 years
of collecting flowering plants in africa. \emph{Global Ecology and
Biogeography}, \emph{25}(9), 1085--1096.

\leavevmode\hypertarget{ref-tank2008annuals}{}%
Tank, D. C., \& Olmstead, R. G. (2008). From annuals to perennials:
Phylogeny of subtribe castillejinae (orobanchaceae). \emph{American
Journal of Botany}, \emph{95}(5), 608--625.

\leavevmode\hypertarget{ref-vaclavik2017effects}{}%
Václavı'k, T., Beckmann, M., Cord, A. F., \& Bindewald, A. M. (2017).
Effects of uv-b radiation on leaf hair traits of invasive
plants---combining historical herbarium records with novel remote
sensing data. \emph{PloS One}, \emph{12}(4), e0175671.

\leavevmode\hypertarget{ref-R-MASS}{}%
Venables, W. N., \& Ripley, B. D. (2002). \emph{Modern applied
statistics with s} (Fourth.). New York: Springer. Retrieved from
\url{http://www.stats.ox.ac.uk/pub/MASS4}

\leavevmode\hypertarget{ref-verboom2004testing}{}%
Verboom, G. A., Linder, H. P., \& Stock, W. D. (2004). Testing the
adaptive nature of radiation: Growth form and life history divergence in
the african grass genus ehrharta (poaceae: Ehrhartoideae).
\emph{American Journal of Botany}, \emph{91}(9), 1364--1370.

\leavevmode\hypertarget{ref-R-ggplot2}{}%
Wickham, H. (2016). \emph{Ggplot2: Elegant graphics for data analysis}.
Springer-Verlag New York. Retrieved from \url{http://ggplot2.org}

\leavevmode\hypertarget{ref-R-tidyverse}{}%
Wickham, H. (2017). \emph{Tidyverse: Easily install and load the
'tidyverse'}. Retrieved from
\url{https://CRAN.R-project.org/package=tidyverse}

\leavevmode\hypertarget{ref-R-forcats}{}%
Wickham, H. (2018a). \emph{Forcats: Tools for working with categorical
variables (factors)}. Retrieved from
\url{https://CRAN.R-project.org/package=forcats}

\leavevmode\hypertarget{ref-R-stringr}{}%
Wickham, H. (2018b). \emph{Stringr: Simple, consistent wrappers for
common string operations}. Retrieved from
\url{https://CRAN.R-project.org/package=stringr}

\leavevmode\hypertarget{ref-R-dplyr}{}%
Wickham, H., François, R., Henry, L., \& Müller, K. (2018). \emph{Dplyr:
A grammar of data manipulation}. Retrieved from
\url{https://CRAN.R-project.org/package=dplyr}

\leavevmode\hypertarget{ref-R-tidyr}{}%
Wickham, H., \& Henry, L. (2018). \emph{Tidyr: Easily tidy data with
'spread()' and 'gather()' functions}. Retrieved from
\url{https://CRAN.R-project.org/package=tidyr}

\leavevmode\hypertarget{ref-R-readr}{}%
Wickham, H., Hester, J., \& Francois, R. (2017). \emph{Readr: Read
rectangular text data}. Retrieved from
\url{https://CRAN.R-project.org/package=readr}

\leavevmode\hypertarget{ref-willis2017old}{}%
Willis, C. G., Ellwood, E. R., Primack, R. B., Davis, C. C., Pearson, K.
D., Gallinat, A. S., \ldots{} others. (2017). Old plants, new tricks:
Phenological research using herbarium specimens. \emph{Trends in Ecology
\& Evolution}, \emph{32}(7), 531--546.

\leavevmode\hypertarget{ref-wolf2016altitudinal}{}%
Wolf, A., Zimmerman, N. B., Anderegg, W. R., Busby, P. E., \&
Christensen, J. (2016). Altitudinal shifts of the native and introduced
flora of c alifornia in the context of 20th-century warming.
\emph{Global Ecology and Biogeography}, \emph{25}(4), 418--429.

\endgroup

\newpage
\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}

\hypertarget{supplement}{%
\section{Supplement}\label{supplement}}

\hypertarget{images-used}{%
\subsubsection{Images used}\label{images-used}}

\url{https://www.gbif.org/occurrence/1099023487}
\url{https://www.gbif.org/occurrence/1057389408}
\url{https://www.gbif.org/occurrence/1099023562}
\url{https://www.gbif.org/occurrence/1099023490}

\hypertarget{software-used}{%
\subsubsection{Software used}\label{software-used}}

We used R (Version 3.5.1; R Core Team, 2018) and the R-packages
\emph{ape} (Version 5.2; Paradis \& Schliep, 2018; Orme et al., 2018;
Soetaert, 2018), \emph{bindrcpp} (Version 0.2.2; Müller, 2018),
\emph{caper} (Version 1.0.1; Orme et al., 2018), \emph{coda} (Version
0.19.2; Plummer, Best, Cowles, \& Vines, 2006), \emph{diagram} (Version
1.6.4; Soetaert, 2017), \emph{dplyr} (Version 0.7.8; Wickham et al.,
2018), \emph{forcats} (Version 0.3.0; Wickham, 2018a), \emph{gee}
(Version 4.13.19; R by Thomas Lumley \& author., 2015), \emph{geiger}
(Version 2.0.6; Alfaro et al., 2009; Eastman, Alfaro, Joyce, Hipp, \&
Harmon, 2011; Harmon, Weir, Brock, Glor, \& Challenger, 2008; Slater et
al., 2012), \emph{ggplot2} (Version 3.1.0; Wickham, 2016),
\emph{logistf} (Version 1.23; Heinze \& Ploner, 2018), \emph{maps}
(Version 3.3.0; Richard A. Becker, Ray Brownrigg. Enhancements by Thomas
P Minka, \& Deckmyn., 2018), \emph{MASS} (Version 7.3.51.1; Venables \&
Ripley, 2002), \emph{Matrix} (Version 1.2.15; Bates \& Maechler, 2018),
\emph{MCMCglmm} (Version 2.26; Hadfield, 2010), \emph{mvtnorm} (Version
1.0.8; Genz \& Bretz, 2009), \emph{papaja} (Version 0.1.0.9842; Aust \&
Barth, 2018), \emph{phylolm} (Version 2.6; Ho \& Ane, 2014),
\emph{phytools} (Version 0.6.60; Revell, 2012), \emph{purrr} (Version
0.2.5; Henry \& Wickham, 2018), \emph{raster} (Version 2.8.4; Hijmans,
2018), \emph{readr} (Version 1.2.1; Wickham et al., 2017), \emph{shape}
(Version 1.4.4; Soetaert, 2018), \emph{sp} (Version 1.3.1; Pebesma \&
Bivand, 2005), \emph{stringr} (Version 1.3.1; Wickham, 2018b),
\emph{tibble} (Version 1.4.2; Müller \& Wickham, 2018), \emph{tidyr}
(Version 0.8.2; Wickham \& Henry, 2018), and \emph{tidyverse} (Version
1.2.1; Wickham, 2017) for all our analyses.

\hypertarget{supplementary-tables-and-figures}{%
\subsubsection{Supplementary tables and
figures}\label{supplementary-tables-and-figures}}

\begin{center}
\begin{ThreePartTable}
\begin{TableNotes}[para]
\normalsize{\textit{Note.} LH = Life history (a = annual, p = perennial). n=sample size of GBIF records. Seasons are mean drought frequencies observed at locations of records.}
\end{TableNotes}
\small{
\begin{longtable}{lllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:speciesmeanstable}Heliophila species records and the mean drought frequencies during different seasons at the location of records }\\
\toprule
Species & \multicolumn{1}{c}{LH} & \multicolumn{1}{c}{n} & \multicolumn{1}{c}{Winter} & \multicolumn{1}{c}{Spring} & \multicolumn{1}{c}{Summer} & \multicolumn{1}{c}{Fall}\\
\midrule
Heliophila acuminata & a & 28 & 0.32 & 0.38 & 0.41 & 0.36\\
Heliophila africana & a & 91 & 0.33 & 0.35 & 0.34 & 0.34\\
Heliophila amplexicaulis & a & 60 & 0.32 & 0.36 & 0.39 & 0.33\\
Heliophila arenaria & a & 65 & 0.34 & 0.37 & 0.38 & 0.34\\
Heliophila carnosa & p & 129 & 0.33 & 0.37 & 0.39 & 0.31\\
Heliophila cedarbergensis & p & 3 & 0.40 & 0.43 & 0.32 & 0.27\\
Heliophila collina & a & 16 & 0.35 & 0.47 & 0.48 & 0.45\\
Heliophila cornellsbergia & a & 2 & 0.33 & 0.42 & 0.35 & 0.21\\
Heliophila cornuta & p & 101 & 0.35 & 0.40 & 0.40 & 0.34\\
Heliophila coronopifolia & a & 40 & 0.37 & 0.42 & 0.40 & 0.37\\
Heliophila crithmifolia & a & 97 & 0.35 & 0.42 & 0.45 & 0.38\\
Heliophila descurva & a & 12 & 0.36 & 0.38 & 0.38 & 0.29\\
Heliophila deserticola & a & 133 & 0.48 & 0.48 & 0.46 & 0.45\\
Heliophila digitata & a & 30 & 0.33 & 0.38 & 0.44 & 0.38\\
Heliophila dregeana & p & 17 & 0.33 & 0.37 & 0.33 & 0.32\\
Heliophila elongata & p & 82 & 0.26 & 0.32 & 0.30 & 0.25\\
Heliophila ephemera & a & 3 & 0.14 & 0.27 & 0.31 & 0.26\\
Heliophila esterhuyseniae & p & 3 & 0.21 & 0.30 & 0.37 & 0.27\\
Heliophila eximia & p & 12 & 0.42 & 0.41 & 0.32 & 0.34\\
Heliophila gariepina & a & 12 & 0.50 & 0.53 & 0.48 & 0.41\\
Heliophila glauca & p & 35 & 0.29 & 0.35 & 0.34 & 0.33\\
Heliophila juncea & p & 150 & 0.32 & 0.37 & 0.39 & 0.35\\
Heliophila linearis & p & 94 & 0.32 & 0.33 & 0.28 & 0.30\\
Heliophila macowaniana & a & 31 & 0.33 & 0.38 & 0.44 & 0.39\\
Heliophila macra & p & 22 & 0.30 & 0.30 & 0.32 & 0.29\\
Heliophila macrosperma & p & 5 & 0.28 & 0.36 & 0.35 & 0.25\\
Heliophila minima & p & 35 & 0.36 & 0.45 & 0.51 & 0.39\\
Heliophila namaquana & a & 16 & 0.39 & 0.46 & 0.48 & 0.39\\
Heliophila nubigena & p & 19 & 0.31 & 0.36 & 0.43 & 0.38\\
Heliophila pectinata & a & 16 & 0.27 & 0.34 & 0.50 & 0.34\\
Heliophila polygaloides & p & 12 & 0.40 & 0.48 & 0.42 & 0.34\\
Heliophila pubescens & a & 9 & 0.31 & 0.40 & 0.48 & 0.39\\
Heliophila pusilla & a & 45 & 0.32 & 0.38 & 0.38 & 0.34\\
Heliophila rigidiuscula & p & 201 & 0.30 & 0.33 & 0.28 & 0.24\\
Heliophila scoparia & p & 106 & 0.31 & 0.37 & 0.36 & 0.31\\
Heliophila seselifolia & a & 80 & 0.36 & 0.42 & 0.45 & 0.40\\
Heliophila suavissima & p & 92 & 0.30 & 0.39 & 0.42 & 0.31\\
Heliophila subulata & p & 103 & 0.29 & 0.33 & 0.31 & 0.29\\
Heliophila tricuspidata & p & 8 & 0.28 & 0.33 & 0.38 & 0.30\\
Heliophila trifurca & a & 77 & 0.45 & 0.48 & 0.48 & 0.43\\
Heliophila tulbaghensis & p & 3 & 0.36 & 0.41 & 0.36 & 0.35\\
Heliophila variabilis & a & 97 & 0.35 & 0.41 & 0.40 & 0.37\\
\bottomrule
\addlinespace
\insertTableNotes
\end{longtable}
}
\end{ThreePartTable}
\end{center}

\begin{figure}
\centering
\includegraphics{../figures/speciesmaps.pdf}
\caption{\label{fig:speciesmaps}Maps of occurance records for individual species.
Orange points indicate annual species. Blue points indicate perennial
species.}
\end{figure}

\end{document}