### 2018-11-28 manuscript/manuscript.Rmd
edits from Kathryn

---
title             : "Drought frequency predicts plant life history strategies"
shorttitle        : "Drought and life history"

author: 
  - name          : "J. Grey Monroe"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "307 University Ave, Fort Collins, CO 80523"
    email         : "monroejg@colostate.edu"
  - name          : "Brian Gill"
    affiliation   : "3"
  - name          : "Kathryn G. Turner"
    affiliation   : "4"
  - name          : "John K. McKay"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO 80523, USA"
  - id            : "2"
    institution   : "College of Agriculture, Colorado State University, Fort Collins, CO 80523, USA"
  - id            : "3"
    institution   : "Institute for Environment and Society, Brown University, Providence, RI 02912, USA"
  - id            : "4"
    institution   : "Biology Department, Pennsylvania State University, State College, PA 16802, USA"

abstract: Explaining variation in life history strategies is a long-standing goal of evolutionary biology. For plants, annual and perennial life histories are thought to reflect adaptation to environments that differ in the frequency of environmental stress such as drought. Here we test this hypothesis in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering plants native to Africa by integrating 2192 herbaria occurrence records with 34 years of satellite-based drought detection. Consistent with predictions from classic life history theory, we find that perennial \textit{Heliophila} species occur in environments where droughts are significantly less frequent compared to annuals. These associations are predictive while controlling for phylogeny, lending support to the hypothesis that drought related natural selection has influenced the distributions of these strategies. Additionally, the difference in drought frequency between annual and perennial species distributions is greatest during the summer and fall, which also appears to be when annuals are in the seed phase of their life cycle based on collection dates of annual species. Together, these finding provide empirical support for classic hypotheses about the drivers of life history strategy in plants - that perennials out compete annuals in environments with less frequent drought and that annuals are adpated to enviroments with more frequent drought by escaping drought prone seasons as seeds.

keywords          : "drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records"

bibliography      : ["r-references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
figsintext        : yes

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf

header-includes   : 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

---

```{r setup, include = FALSE}
library("papaja")
library(raster)
library(tidyverse)
library(ggplot2)
library(logistf)
library(geiger) 
library(ape) 
library(phylolm)
library(gridExtra)
library(cowplot)
library(rasterVis)
library(ggtree)
library(grid)
library(png)
```

# Introduction

Understanding the causes and consequences of life history variation is a longstanding goal of ecology and evolutionary biology  [@cole1954population]. In plants, life histories are especially diverse, with herbaceous species that complete their life cycle in a number of weeks to trees that live for thousands of years [@brown1996oldlist]. Along this continuum an important division exists, distinguishing annuals which complete their seed to seed life cycle within a single calendar year from perennials which can persist over multiple years. Annual plants flower once, set seed, senesce, and then die, spending at least some portion of the year as a seed, where they are relatively protected from environmental stress. In contrast, perennial plants can continue vegetative growth after reproduction and must survive conditions experienced during all seasons. These represent fundamentally different life history strategies, but the ecological factors that explain their evolution and distributions remain empirically uresolved [@friedman2015all].  
Classical theory predict shorter life spans in environments where adult mortality is high [@charnov1973life; @stearns1992evolution;@franco1996life]. In plants, this has been extended to the hypothessis that annuality is adaptive when it allows plants to escape drought [@schaffer1975selection]. Lack of water is perhaps the greatest threat to survival during vegetatitive or reproductive growth and annuals can remain dormant (and protected as a seed) during drought. Thus, environments with greater seasonal drought frequency may select for annual life histories that complete reproduction prior to drought prone seasons. Conversely, environments with less frequent drought may select for perennial species, which may benefit from multiple bouts of reproduction and competitive advantage by preventing recruitment of annual species [@corbin2004competition]. These predictions have been supported by the association of annualilty with arid environments in _Oryza perennis_ [@morishima1984differentiation] and _Oenothera_ [@evans2005climate]. Additionally, annual and perennial species of _Nemesia_ were qualitatively associated with winter rather and summer rainfall environments respectively [@datson2008climate] and annual species of _Scorzoneroides_ were associated with environments classified as unpredictable [@cruz2009molecular]. However, whether the history frequency of drought events indeed predicts the distributions annual or perennial life history strategies has yet to be tested in a phylogenetic context with suitable sample sizes of distrubution and climate data.  
Here we combine a long-term global dataset of satellite detected drought events with metadata from natural history collections to test these classic hypotheses about the evolution of life history strategies within the African endemic mustard genus, _Heliophila_ L. (Brassicaceae). If annuallity is an adaptive strategy allowing plants to escape drought prone seasons, then drought frequency should predict the distribution of life history strategies across landscapes, and annual species should be more commonly associated with drought prone regions than perennial species. Furthermore, if annual species have adapted to escape drought prone seasons, observations of growing annual species (i.e. occurring in forms other than seed) should be rare during drought prone seasons. Phylogenetic relatedness can have significant non-random effects on species distributions and life history traits [@barrett1996comparative], and therefore we assessed the relationship between life history distribution and drought frequency in a phylogenetically controlled background.

# Materials and Methods
## Data  
### Availability
All analyses were peformed using R. All data and the source code to produce this manuscript are available at https://github.com/greymonroe/heliophila. Softare used is listed in the supplement.

### Satellite-detected drought data
```{r Droughtdata}
Winter<-raster("../data/drought/Summer_drought_freq")
Spring<-raster("../data/drought/Fall_drought_freq")
Summer<-raster("../data/drought/Winter_drought_freq")
Fall<-raster("../data/drought/Spring_drought_freq")
```
Remotely sensed data is a powerful tool for characterizing seasonal patterns in drought because it is less limited in spatial and temporal scope and resolution than weather stations or field observations [@aghakouchak2015remote]. To quantify the frequency of drought during different seasons across landscapes, we used the remotely sensed Vegetative Health Index (VHI), which measures landscape scale reductions in plant cover and temperature conditions characteristic of drought [@kogan2001operational]. Generated from data collected by NOAA AVHRR satellites since 1981, the VHI combines Normalized Difference Vegetation Index (NDVI) derived measures of vegetative stress (Vegetative Condition Index - VCI) with temperature stress indicated by anomalies in thermal spectra (Temperature Condition Index - TCI). The VHI of year $y$ during week $w$ of $[1,52]$ at pixel $i$ is derived from the following equations, where $n$ is the number of years observed.

$$VCI_{y,w,i} = 100\frac{NDVI_{y,w,i} - NDVI_{min}}{NDVI_{max} - NDVI_{min}}$$

$$TCI_{y,w,i} = 100\frac{T_{y,w,i} - T_{min}}{T_{max} - T_{min}}$$

$$VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})$$
where $NDVI_{min} = min(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})$ and $NDVI_{max} = max(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})$ and $T_{min} = min(T_{1981,w,i}...T_{1981+n,w,i})$ and $T_{max} = max(T_{1981,w,i}...T_{1981+n,w,i})$

Thus, VHI measurements are standardized according to conditions historically observed at each locations. These measurements have been validated and generally used for evaluating drought risk and predicting crop yields in agriculture [e.g., @rojas2011assessing; @kogan2016modelling]. But they also present a new tool to study seasonal patterns in the frequency of drought across environments and to test hypotheses about the effect of drought on ecological and evolutionary processes [@kerr2003space]. As such, the VHI  has been applied recently to study drought related ecology of natural species and proven useful for predicting infraspecific variation in drought tolerance traits and genes [@Mojica2016;@dittberner2018natural;@monroe2018drought]. Here, we accessed VHI data at $16km^2$ resolution from 1981 to 2015 (https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php) to characterize the seasonal drought frequencies experienced by annual and perennial _Heliophila_ species.

### Life history data for Heliophila
_Heliophila_ is a genus of flowering plants endemic to the southern portion of Africa including the Cape Floristic and Succulent Karoo Regions. These are among the most botanically diverse environments on Earth and the estimated ~50 _Heliophila_ species occurring there are considered to make up the most diverse genus of the family Brassicaceae [@mummenhoff2005phylogeny;@mandakova2012whole]. This genus includes both perennial and annual species and this change in life history strategy has likely arisen multiple independent times [@appel1997generic;@mummenhoff2005phylogeny]. Furthermore, the fine scale climatic heterogeneity of Southern Africa is ideal for studying the distribution of traits in relation to environmental parameters [@sayre2013new]. We used life histories reported by @mummenhoff2005phylogeny, grouping species with annual or perennial life histories. Perenniality was defined based any form of perennial life history (e.g., herbs, shrubs, mixed, etc). We also repeated these analyses without _H. subulata_ and _H. minima_ which have mixed life histories and found similar results. Because the original authors didn't report the nature of the mixed trait [@mummenhoff2005phylogeny], (ie. plasticity vs. genetic variation), we classified these species here as perennial since they have the capacity to maintain vegetative growth after reproduction at least to some capacity.

### Heliophila occurance records
```{r raw_GBIF}
GBIF<-read.csv("../data/0006178-160311141623029/occurrence.csv")
```
Botanists have collected and maintained over 350 million botanical specimens worldwide over the past 300 years. Herbarium specimens and their associated metadata have been used since the 1960s to study species’ geographical distributions (reviewed by  @willis2017old and @lang2018using). And as they become digitized [@soltis2017digitization], these collections have been used to study relationships between trait distributions, geography, and climate [@davis2015herbarium;@wolf2016altitudinal;@stropp2016mapping;@vaclavik2017effects]. To characterize the diributions of annual and perennial _Heliophila_ species, all records for the genus _Heliophila_ were downloaded from the Global Biodiversity Information Facility (gbif.org) on July 21, 2018 [@gbifdownload].

### Sequence data for phylogeny
Aligned Heliophila ITS sequences were obtained from previous work by @mandakova2012whole.  Aethionema, Alliaria, Cardamine, Chamira, and Rorippa ITS records from were downloaded from Genbank.

## Analyses
### Drought frequency calculations
To characterize drought regimens across the distrubtions of annual and perennial species of _Heliohpila_, we calculated drought during different seasons at the location of observations for _Heliophila_ records using the VHI. Specifically, we created global maps of the frequencies of observing drought conditions (VHI<40, NOAA) during the winter (quarter surrounding winter solstice), spring (quarter surrounding spring equinox), summer (quarter surrounding summer solstice) and fall (quarter surrounding fall equinox) from 1981 to 2015. From these maps, the drought frequency during the winter, spring, summer, and fall were extracted for the locations of all GBIF records.

### Filtering of occurance records
```{r cleanup}
species_life_history<-read.csv("../data/Heliophila_life_history.csv")
species_life_history$species<-gsub("_", " ", species_life_history$species)
species_life_history$species <- as.factor(species_life_history$species)

GBIF <- GBIF %>%
  mutate(Winter=raster::extract(Winter, cbind(decimalLongitude, decimalLatitude))) %>%
  mutate(Spring=raster::extract(Spring, cbind(decimalLongitude, decimalLatitude))) %>%
  mutate(Summer=raster::extract(Summer, cbind(decimalLongitude, decimalLatitude))) %>%
  mutate(Fall=raster::extract(Fall, cbind(decimalLongitude, decimalLatitude))) %>%
  mutate(ymd_lat_lon_sp = paste0(year,"_",month,"_",day,"_", decimalLatitude,"_", decimalLongitude,"_", species))

GBIF_cleaned <- GBIF %>%
  filter(species %in% species_life_history$species) %>% 
  filter(!is.na(decimalLatitude)) %>%
  filter(hasGeospatialIssues==FALSE) %>%
  filter(decimalLatitude<(-20)) %>%
  filter(decimalLongitude<40) %>%
  filter(!duplicated(ymd_lat_lon_sp)) %>%
  merge(species_life_history) %>%
  filter(!is.na(Summer))
```

To avoid instances with spurious location data, we filtered raw GBIF by restricting our analyses to include only:

* records for species with reported life history  
* records with geospatial data  
* records without known geospatial coodinate issues (i.e., coordinates reported are those of herbarium)  
* records from collection sites classified as land pixels   
* records from Africa (to exclude locations of cultivation)
* records without duplicates (i.e., identical species, location, collection date)

### Phylogeny construction
Out group species were aligned together with _Heliophila_ ITS sequences using MAFFT. Model selection for construction of phylogeny was performed in jModeltest2 with CIPRES. Based on this analysis, $GTR + L$ were selected. Ultrametric phylogeny was estimated with branch lengths as relative time.

### Comparison of drought frequency between annual and perennial species
```{r speciesmeans}
species_means <- GBIF_cleaned %>%
  dplyr::select(species, life_history, Winter, Summer, Spring, Fall) %>%
  group_by(species, life_history) %>%
  summarise(n=n(), Winter_mean=mean(Winter), Spring_mean=mean(Spring), Summer_mean=mean(Summer), Fall_mean=mean(Fall)) %>% 
  as.data.frame()

```
```{r firthmodels, echo=F,  results="hide"}

firthmodelstable<-data.frame()

# non-phylogenetic logistic regression firth penalized
species_means$life_history_num<-as.numeric(species_means$life_history)-1
row.names(species_means)<-gsub(" ", "_", species_means$species)
seasons<-c("Winter_mean", "Spring_mean", "Summer_mean","Fall_mean")
for(s in 1:length(seasons)){
  fit1 <- logistf(species_means$life_history_num~(species_means[,seasons[s]]), firth = TRUE)
  sum<-summary(fit1)
  row.names(sum$coefficients)<-NULL
  sum_df<-data.frame(Predictor=c("Intercept", gsub("_mean", " drought freq.", seasons[s])), 
                     Estimate=sum$coefficients, 
                     p.value=sum$prob)
  row.names(sum_df)<-NULL
  firthmodelstable<-rbind(firthmodelstable, sum_df)
}

```
```{r phylomodels, results="hide"}
# phylogentic constraint
tree<-read.nexus("../data/TrimmedMCCwoPolytomies.nex")
tree<- rescale(tree, "depth", 1)
tree<-keep.tip(tree, tip = gsub(" ","_", as.character(species_means$species)))
phylomodelstable<-data.frame()

for(s in 1:length(seasons)){
  fit1 <-  phyloglm(species_means$life_history_num~species_means[,seasons[s]], data=species_means, phy=tree)
  sum<-summary(fit1)
  row.names(sum$coefficients)<-NULL
  sum_df<-data.frame(Predictor=c("Intercept", gsub("_mean", " drought freq.", seasons[s])), sum$coefficients)
  phylomodelstable<-rbind(phylomodelstable, sum_df)
}

phylomodelstable <- phylomodelstable[,c( "Predictor","Estimate","p.value")]

```
To evaluate the hypothesis that annual and perennial life history strategies reflect adaptations to alternative drought regimes, we tested the corresponding prediction that the observed distributions of annual and perennial _Heliophila_ species would be significantly associated with historic drought frequency.  First, we compared the frequency of drought during the winter, spring, summer, and fall between raw occurance records of annual and perennial species by t-tests. To account for variation in the number of occurance records per species, we next calculated the mean drought frequency during the winter, spring, summer and fall for each species. The relationships between species mean values of drought frequency during each season and life habitat (annual or perennial) were tested using Firth's penalized-likelihood logistic regressions.  Because demographic histories caused by ancestry can confound trait - environment associations, we then tested for the relationships between drought frequeny and life history while controlling for relationships between species using phylogenetic logistic regressions.

### Collection dates

To test the hypothesis that annual species have adapted to escape drought prone seasons as seeds, collection dates for herbarium specimens were compared between annual and perennial species. Comparisons of distributions were made by Two-sample Kolmogorov-Smirnov test, t-test, and Barlett variance test.

# Results
(ref:phylogeny) Species and examples of herbaria specimens of _Heliophila_ (a) Phylogeny and life history strategies of species studied. Orange circles at branch tips mark annual species and blue circles mark perennial species. Example herbaria specimens accessed via GBIF of (a) _H. minima_, (b) _H. deserticola_, (c) _H. coronopifolia_ and (d) _H. ephemera_. Images (a,c,d) courtesy of The Bavarian Natural History Collections (CC BY-SA 4.0) and (b) The London  Natural History Museum (CC BY 4.0). Links to images are found in the supplement.

```{r makephylogeny, eval=T, message=F, warning="hide", results="hide"}

pdf("../figures/phylogeny.pdf", height=6, width=5)

tree<-read.nexus("../data/TrimmedMCCwoPolytomies.nex")
tree<- rescale(tree, "depth", 1)
tree$tip.label<-gsub("_"," ", tree$tip.label)
tree<-keep.tip(tree, tip =as.character(species_means$species))

treeplot<-ggtree(tree, layout = "rectangular")+
  xlim(c(0,2))+
  geom_tippoint(size=3,col=ifelse(species_life_history$life_history[match(tree$tip.label,species_life_history$species)]=="a", "orange3", "dodgerblue3"))+
  geom_tiplab(offset = .05 , size=3)+
  theme(plot.background = element_rect(fill=alpha("white",0)))

minima <- readPNG("../pictures/minima.png")
minima <- rasterGrob(minima, interpolate=TRUE)
#https://www.gbif.org/occurrence/1099023487

deserticola <- readPNG("../pictures/deserticola.png")
deserticola <- rasterGrob(deserticola, interpolate=TRUE)
#https://www.gbif.org/occurrence/1057389408

coronopifolia<-readPNG("../pictures/coronopifolia.png")
coronopifolia <- rasterGrob(coronopifolia, interpolate=TRUE)
#https://www.gbif.org/occurrence/1099023562

ephemera<-readPNG("../pictures/ephemera.png")
ephemera <- rasterGrob(ephemera, interpolate=TRUE)
#https://www.gbif.org/occurrence/1099023490

pics<-plot_grid(minima, deserticola, coronopifolia, ephemera,
                labels = c("b","c","d","e"), ncol = 1, hjust = -1)

plot_grid(treeplot, pics, ncol=2,labels=c("a",""), rel_widths = c(1,0.5))

dev.off()

```
```{r phylogeny, fig.cap='(ref:phylogeny)', out.width = "\\textwidth", fig.pos = "!h"}
knitr::include_graphics("../figures/phylogeny.pdf", dpi = 108)
```
(ref:mapsboxplots) Locations of (a) annual and (b) perennial Heliophila. Drought frequency during the (c) winter, (d) spring, (e) summer and (f) fall. (g) Drought frequencies during each season observed at the collection sites of _Heliophila_ records.

```{r make_maps_boxplots, eval=T, message=F, results="hide"}

pdf("../figures/maps_boxplots.pdf", width=5, height=7.5)
world <- map_data("world") # we already did this, but we can do it again

perennialmap<-ggplot() + geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
  geom_point(GBIF_cleaned %>% filter(life_history=="p"), mapping=aes(x=decimalLongitude, y=decimalLatitude), pch=3, col="dodgerblue3", alpha=0.3) +
  theme_classic()+
  theme( text = element_text(size=8))+
  labs(title="Perennial species", x="Longitude (°)", y="Latitude (°)")+
  annotate("text", x=32, y=-34, label=paste("n obs. =", nrow(GBIF_cleaned %>% filter(life_history=="p"))), size=2)

annualmap<-ggplot() + geom_polygon(data = world, aes(x=long, y = lat, group=group),  lwd=.5, fill = NA, color = "black") + 
  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
  geom_point(GBIF_cleaned %>% filter(life_history=="a"), mapping=aes(x=decimalLongitude, y=decimalLatitude), pch=3, col="orange3", alpha=0.3) +
  theme_classic()+
  theme(text = element_text(size=8))+
  labs(title="Annual species", x="Longitude (°)", y="Latitude (°)")+
  annotate("text", x=32, y=-34, label=paste("n obs. =", nrow(GBIF_cleaned %>% filter(life_history=="a"))), size=2)

Winter_SA <-crop(Winter, rbind(c(12,37),c(-35, -18)))

wintermap<-gplot(Winter_SA) + 
  geom_tile(aes(fill=value)) + 
  scale_fill_gradientn(colors=c( 'white',"gray80", "gray20", 'black'), na.value="white", name=NULL) +
  geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
  theme_classic()+
  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.5,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
  labs(title="Winter drought freq.", x="Longitude (°)", y="Latitude (°)")

Spring_SA <-crop(Spring, rbind(c(12,37),c(-35, -18)))

springmap<-gplot(Spring_SA) + 
  geom_tile(aes(fill=value)) + 
  scale_fill_gradientn(colors=c( 'white',"gray80", "gray20", 'black'), na.value="white", name=NULL) +
  geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
  theme_classic()+
  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.5,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
  labs(title="Spring drought freq.", x="Longitude (°)", y="Latitude (°)")

Summer_SA <-crop(Summer, rbind(c(12,37),c(-35, -18)))

summermap<-gplot(Summer_SA) + 
  geom_tile(aes(fill=value)) + 
  scale_fill_gradientn(colors=c( 'white',"gray80", "gray20", 'black'), na.value="white", name=NULL) +
  geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
  theme_classic()+
  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.5,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
  labs(title="Summer drought freq.", x="Longitude (°)", y="Latitude (°)")

Fall_SA <-crop(Fall, rbind(c(12,37),c(-35, -18)))

fallmap<-gplot(Fall_SA) + 
  geom_tile(aes(fill=value)) + 
  scale_fill_gradientn(colors=c( 'white',"gray80", "gray20", 'black'), na.value="white", name=NULL) +
  geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
  theme_classic()+
  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.5,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
  labs(title="Fall drought freq.", x="Longitude (°)", y="Latitude (°)")

GBIF_cleaned_long<- GBIF_cleaned %>%
  dplyr::select(species, life_history, Winter, Spring, Summer,Fall) %>%
  gather(season, drought_freq, Winter, Spring, Summer,Fall)

GBIF_cleaned_long$season<-factor(GBIF_cleaned_long$season, levels=c("Winter","Spring","Summer","Fall"))

boxplots<-ggplot(GBIF_cleaned_long, aes(x=life_history, col=life_history, y=drought_freq))+
  facet_wrap(~season, ncol=4 )+
  geom_jitter(width = .3, alpha=0.1, pch=3)+
  geom_boxplot(outlier.colour = "white", outlier.fill = NA, fill=NA, width=.3)+
  scale_color_manual(values=c("orange3","dodgerblue3"))+
  theme_classic()+
  theme(legend.position = "none", text = element_text(size=8))+
  scale_x_discrete(name="Life history", labels=c("Annual","Perennial"))+
  labs(y="Drought frequency")

maps<-plot_grid(annualmap, 
          perennialmap, 
          wintermap, 
          springmap,
          summermap,
          fallmap, 
          labels = "auto", ncol = 2)
plot_grid(maps, boxplots, ncol=1,labels=c("","g"), rel_heights = c(.75,.35))

dev.off()

```

```{r mapsboxplots, fig.cap='(ref:mapsboxplots)', out.width = "\\textwidth", fig.pos = "!h"}
knitr::include_graphics("../figures/maps_boxplots.pdf", dpi = 108)
```
Out of `r nrow(GBIF)` _Heliophila_ GBIF recrods,  `r sum(GBIF$species %in% species_life_history$species)` were for species with reported life history [@mummenhoff2005phylogeny], `r sum(!is.na(GBIF$decimalLatitude))` had geospatial data, `r sum(!(GBIF$hasGeospatialIssues==T) & !(is.na(GBIF$decimalLatitude)))` did not have geospatial issues, `r sum(!(is.na(GBIF$Summer)) & !(is.na(GBIF$decimalLatitude)))` were located on pixels classified as land having drought measurements, `r sum(GBIF$decimalLatitude < (-20) & GBIF$decimalLongitude < (40) & !is.na(GBIF$decimalLatitude))` were located in Africa, `r sum(!duplicated(GBIF$ymd_lat_lon_sp) & !is.na(GBIF$decimalLatitude)) ` were not duplicated. After all filtering steps, `r nrow(GBIF_cleaned)` records for `r length(unique(GBIF_cleaned$species))` species (Figure \@ref(fig:phylogeny), Table \@ref(tab:speciesmeanstable)) passed for further analyses. The number of samples varied between species, with a mean of `r mean(species_means$n)` samples per species.  `r as.character(species_means$species[species_means$n==max(species_means$n)])` had the most records, `r max(species_means$n)`,  and  `r as.character(species_means$species[species_means$n==min(species_means$n)])` the fewest,  `r min(species_means$n)` (Table \@ref(tab:speciesmeanstable)).

There were clear visual differences between the distributions of the `r sum(GBIF_cleaned$life_history=="a")` annual and the `r sum(GBIF_cleaned$life_history=="p")` perennial _Heliophila_ observation records (see Figure \@ref(fig:speciesmaps) for maps of individual species). While annual species were generally found in the western regions of South Africa and Namibia, primarily in the Cape Floristic Region and Succulent Karoo (Figure \@ref(fig:mapsboxplots)a), the occurances of perennials extended to the east coast of South Africa (Figure \@ref(fig:mapsboxplots)b).

```{r ttests, eval=T, message=F, warning="hide", results="hide"}
twin<-t.test(Winter~life_history, GBIF_cleaned)
tspr<-t.test(Spring~life_history, GBIF_cleaned)
tsum<-t.test(Summer~life_history, GBIF_cleaned)
tfal<-t.test(Fall~life_history, GBIF_cleaned)
```

The frequency of drought varied considerably across the ranges of _Heliophila_ species (Figure \@ref(fig:mapsboxplots)c-f). This heterogeneity is expected, given that this is one of the most climatically diverse regions of the Earth [@sayre2013new]. It is worth noting the east to west cline in drought frequency observed during the summer, which distinguishes the high drought frequency of the Kalahari Sands and Namid Destert phyotogeographic regions from the low drought frequency of the Drakensberg Mountains and Coastal Zambesian phytogeographic regions. In the Cape phytogeographic region there was finer scale heterogeneity in drought frequency during the summer.

Theory predicts that annuality should be adaptive in places where stresses such as drought are more common. Conversely, perenniality should be adaptive in places where such stresses are less frequent. We found that the frequency of drought was significantly higher at the locations of occurance records for annual species. In terms of raw observation records 
(Figure \@ref(fig:mapsboxplots)g), the frequency of drought was signficantly higher at the location of annuals during the winter (t = `r twin$statistic`, p = `r twin$p.value`), spring (t = `r tspr$statistic`, p = `r tspr$p.value`), summer (t = `r tsum$statistic`, p = `r tsum$p.value`), and fall (t = `r tfal$statistic`, p = `r tfal$p.value`). Because raw occurance records do not account for variation in the number of records per species (Table \@ref(tab:speciesmeanstable)), we also tested whether mean drought frequency values of each species were significantly different between annuals and perennials by a Firths-penalized logistic regression. We found that the mean drought frequencies were signficantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a). We further tested whether annual species are found in places where droughts occur more frequently while controlling for the phylogentic relatedness of _Heliophila_ species. This is important, because environmental differences in species distributions can be confounded with demographic history caused by acestry. We found that while controlling for phylogeny, the mean drought frequencies were signficantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a). These findings indicate that common acestry alone does not explain differences the drought frequencies experienced in the environemnts of annual and perennial _Heliophila_.

```{r modelstable, results = 'asis', echo = F}

modelstable<-bind_cols(firthmodelstable,phylomodelstable)

modelstable<- modelstable %>%
  dplyr::select(Predictor, Estimate, p.value,Estimate1, p.value1)

colnames(modelstable)<-c("Predictor", "Estimate`","P`","Estimate*","P*")

apa_table(modelstable
  , caption = "Logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species the winter, spring, summer, and fall."
  , note = "` = Firth's penalized logistic regression. * = Phylogenetically constrained logistic regression. Annual species were scored as 0 and perennial species as 1.",
  midrules=c(2,4,6,8),
  digits=4
)

```

(ref:lineplots) (a) Comparison (mean +- SE) of drought frequency across seasons measured at the GBIF records of annual and perennial species of Heliophila. (b) Collection dates of GBIF records of annual and perennial species of Heliophila.

```{r makelineplots, eval=T, message=F, results="hide", warning=F}

pdf("../figures/line_and_dates.pdf", height=5, width=5)

life_history_drought_means<- species_means %>% 
  dplyr::select(life_history, Winter_mean, Spring_mean, Summer_mean,Fall_mean)  %>%
  gather(variable, value, -life_history) %>%
  group_by(life_history, variable) %>%
  summarise(mean = mean(value), se = sd(value)/sqrt(length(value))) %>%
  mutate(variable = reorder(variable, c(4,2,3,1)))

lines<-ggplot(life_history_drought_means, aes(x=variable, col=life_history,group=life_history, y=mean))+
  theme_classic()+
  geom_point(size=2)+
  geom_line(lwd=0.5, lty=1)+
  geom_errorbar(width=0.25,lwd=1, aes(ymin=mean-se, ymax=mean+se))+
  scale_y_continuous(limits=c(0.3,0.45), labels = c("0.300", "0.350", "0.400", "0.450"))+
  scale_x_discrete(labels=c("Winter","Spring","Summer","Fall"))+
  scale_color_manual(values=c( "orange3","dodgerblue3"), name="Life history", labels=c("Annual", "Prennial"))+
  labs(x="Season", y="Drought Frequency")+
  annotate("text", x=2, y=0.44, label="*", size=7) +
  annotate("text", x=3:4, y=0.44, label="**", size=7)

GBIF_cleaned$date_calc<-as.Date(paste(
  GBIF_cleaned$month,
  GBIF_cleaned$day, sep="-"),"%m-%d")
GBIF_cleaned$days<-lubridate::yday(GBIF_cleaned$date_calc)
GBIF_cleaned$days_adjusted<-sapply(GBIF_cleaned$days, function(x)
  if(is.na(x)){ return(NA)
  }else if(x>135) {
    return (x-135)
  }else return (x+230))

dates<-ggplot(GBIF_cleaned %>% filter(is.finite(days_adjusted)), aes(x=days_adjusted, fill=life_history)) +
  geom_density(alpha=.5)+
  scale_fill_manual(values=c("orange3","dodgerblue3"), name="Life history", labels=c("Annual", "Perennial"))+
  theme_classic()+
  scale_x_continuous(name="Season", breaks=c(45,135,225, 315), labels=c("Winter","Spring","Summer","Fall"))+
  labs(title="Date of observation for occurance records", y="Density")+
  theme(plot.title = element_text(hjust = 0.5, size=4))

plot_grid(lines, dates, ncol=1, labels="auto")

annuals<-subset(GBIF_cleaned, life_history=="a")
perennials<-subset(GBIF_cleaned, life_history=="p")

kout<-ks.test(annuals$days,perennials$days)
tout<-t.test(days~life_history, GBIF_cleaned)
bout<-bartlett.test(days~life_history, GBIF_cleaned)

dev.off()

```

```{r lineplots, fig.cap='(ref:lineplots)'}
knitr::include_graphics("../figures/line_and_dates.pdf", dpi = 108)
```

The preceding results indicated that annual speices are found in environments where droughts are signifacantly more frequent, especially in the summer and fall. Classic life history theory hypothesizes that annuality reflects adaptation to such environments because it allows species to escape stressful conditions. If this is the case, we would expect that annuals spend the drought prone seasons of summer and fall as seeds. To test this hypothesis, we compared the dates of occurance records between annual and perennial _Heliophila_ species. The distributions reveal a considerable difference in the timing of observation of these two life histories. In comparison to perennials, which appear to be collected throughout the year, annuals are almost exclusively observed during the winter and spring (Figure \@ref(fig:lineplots)b). The differences between the distrubution of collection dates were signicant by all tests (ks.test D = `r  kout$statistic`, p = `r  kout$p.value`; bartlett.test K2 = `r bout$statistic`, p = `r bout$p.value`) This is consistent with a model of life history in which annual species flower in the spring, set seed, senesce, and die before the summer. Thus, these annual species are likely to remain dormant during the summer and fall, when drought is the strongest predictor of the distrubutions of annual and perennial life histories (Figure \@ref(fig:lineplots)a).

# Discussion

To test the hypothesis that annual and perennial plants reflect adaptation to alternative drought environments we examined the landscape distribution of life history strategies in the large and diverse mustard genus, _Heliophila_. Using metadata of `r nrow(GBIF_cleaned)` occurance records and a 34 year dataset of satellite-detected droughts, we tested the prediction that annual species are more often observed in drought-prone locations than perennial species, when controlling for phylogenetic relatedness. We found that drought frequency is significantly different between the distributions of annual and perennial species, with annuals being found in environments with more frequent drought, and that this signal is strongest during the seasons when annuals are likely escaping via seed dormancy These results remain significant while controlling for the phylogenetic relationships of _Heliophila_ species, yielding support for the role that natural selection has played in driving contemporary distributions of these alternatives strategies in relation to drought regimens.

We cannot elimiate the possibility that counfounding traits or environmental variables are the causative factors explaining variation in the distributions of annual and perennial species. Nevertheless, these results reveal quantitative support for the classic prediction that annual species are found in environemnts that experience more frequent drought than perennial species. These findings complement previous reports of qualitative associations between annaulity with enviroments characterized as having increased aridity [@evans2005climate], alternative precipitation defined habitats [@morishima1984differentiation;@datson2008climate], or greater unpredictability [@cruz2009molecular]. However, to our knowledge this is the first study to demonstrate a signficant association between life history and drought in a phylogenetic context informed by large scale species distribution data and long term drought measures.

Unfortunately, herbarium collections and their associated data do not represent systematic or random sampling of a species distribution. Significant biases in collecting exist, which we have not necessarily controlled for here, and may have some effect on our findings, such as a bias toward collecting near roads or near the locations of natural history collections [@daru2018widespread]. Future research will benefit from systematic sampling efforts to avoid these noted biases. However, the ecosystems of southern Africa include several biodiversity hotspots and are among the most botanically well sampled regions on Earth [@daru2018widespread], suggesting that this may currently be the optimal region for our analyses of life history distribution. Indeed, we were able to use `r nrow(GBIF_cleaned)` occurance records to study 42 species, which represents a significant advance over relying on personal observations to characterize species distrubtuions.

These findings support classical theoretical predictions about the adaptive value of annual and perennial life history strategies. Taken together, they suggest that in _Heliophila_, annual species are adapted to environments with increased summer droughts by avoiding these seasons in a dormant seed phase of their life cycle. They also suggest that perenniality is adaptive in environemnts where droughts are less frequent. While most previous work has focused on describing the evolutionary origins of annuality [@barrett1996comparative;@conti1999phylogenetic;@andreasen2001unequal;@verboom2004testing;@friedman2015all] there are at least a few other cases where perenniality appears to have arisen from an annual acestor [@bena1998molecular;@tank2008annuals]. And while early theory predicted selection for annuality when adult morality is high [@stearns1992evolution],  we also find evidence that the transition to perenniality could be explained by historical drought regimens. The phylogeny reveals several transitions from annual to perennial life history (Figure \@ref(fig:phylogeny)a) and that the distributions of perennial _Heliophila_ extend into regions where drought frequency is low (Figure \@ref(fig:mapsboxplots)b,Figure \@ref(fig:speciesmaps)).  Perennials may be able to out complete annual relatives in environments where the infrequency of drought favors strategies that allow plants to benefit from growth over many seasons. It may also indicate that annuals rely on drought as a source of disturbance for seedling recruitment when competing with perennials [@corbin2004competition]. Indeed, no annual species were observed in the low drought regions of eastern South Africa \@ref(fig:mapsboxplots)a,Figure \@ref(fig:speciesmaps)).

These findings suggest that species with locally adaptive life history strategies could be threatened by rapidly changing drought regimens [@dai2011drought]. This could have impacts on ecosystem functioning and processes such as carbon cycling if the composition of annual and perennial species changes as a response [@garnier1997specific; @roumet2006suites;monroe2018ecoevolutionary]. Furthermore, the frequency of drought may be an important factor when considering the use of perennial cropping sysyems [@parry2005prospects;@lelievre2009current].

In conclusion, we find strong support for classic life history theory which predicts that annuality is adaptive in environments where droughts occur more frequently. Additionally, we report evidence consistent with a life history model in annuals in which they escape drought prone seasons during the seed phase of their life cycle. Finally, we find evidence that the distributions of perennial lineages may indicate a competitive advantage in areas where droughts are infrequent. More broadly, this work highlights the irreplaceable value of natural history collections and demonstrates the power of combining such information with large scale remote sensing data to address outstanding classic hypotheses in ecology and evolution.

# Acknowledgments 
We thank Jesse Lasky and members of the Sloan lab for generous feedback that improved the quality of this work. This research was supported by NSF Award 1701918 and USDA-NIFA Award 2014-38420-21801 to JGM.

# References
\newpage
```{r create_r-references}
#r_refs(file = "r-references.bib")
```
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

\newpage
\beginsupplement

# Supplement

### Images used
https://www.gbif.org/occurrence/1099023487
https://www.gbif.org/occurrence/1057389408
https://www.gbif.org/occurrence/1099023562
https://www.gbif.org/occurrence/1099023490

### Software used
We used `r cite_r("r-references.bib")` for all our analyses.

### Supplementary tables and figures
```{r speciesmeanstable, results = 'asis', echo = F}
species_means_table<-species_means
row.names(species_means_table) <- NULL
species_means_table<-species_means_table %>% dplyr::select(-life_history_num)
colnames(species_means_table)<-c("Species","LH","n","Winter","Spring","Summer","Fall")
apa_table(
  species_means_table
  , caption = "Heliophila species records and the mean drought frequencies during different seasons at the location of records "
  , note = "LH = Life history (a = annual, p = perennial). n=sample size of GBIF records. Seasons are mean drought frequencies observed at locations of records.",
  digits=2,
  font_size="small",
  longtable = TRUE
)
```

(ref:speciesmaps) Maps of occurance records for individual species. Orange points indicate annual species. Blue points indicate perennial species.

```{r make_spceiesmaps, results = 'hide', echo = F, warnings="hide"}
pdf("../figures/speciesmaps.pdf", height=8, width=8)

world <- map_data("world") # we already did this, but we can do it again

plot_grid(plotlist=lapply(unique(GBIF_cleaned$species), function(x){

  lh<-unique((GBIF_cleaned %>% filter(species==x))$life_history)
  lh_col=ifelse(lh=="a", "orange3", "dodgerblue3")

  ggplot() + geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
    coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
    geom_point(GBIF_cleaned %>% filter(species==x), mapping=aes(x=decimalLongitude, y=decimalLatitude), pch=3, col=lh_col, alpha=0.3) +
    theme_classic()+
    theme( text = element_text(size=5))+
    labs(title=x, x=NULL, y=NULL)

}), ncol=6)

dev.off()

```

```{r speciesmaps, fig.cap='(ref:speciesmaps)'}
knitr::include_graphics("../figures/speciesmaps.pdf", dpi = 108)
```