### 2019-03-08 manuscript/manuscript.tex

\documentclass[man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Drought frequency predicts life history strategies in Heliophila},
            pdfauthor={J. Grey Monroe, Brian Gill, Kathryn G. Turner, \& John K. McKay},
            pdfkeywords={drought adaptation, life history evolution, remote sensing,
phylogeography, herbaria records, Heliophia},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Drought frequency predicts life history strategies in \emph{Heliophila}}
    \author{J. Grey Monroe\textsuperscript{1,2}, Brian Gill\textsuperscript{3},
Kathryn G. Turner\textsuperscript{4}, \& John K.
McKay\textsuperscript{2}}
    \date{}

\shorttitle{Drought and life history}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO 80521, USA\\\textsuperscript{2} College of Agriculture, Colorado State University, Fort Collins, CO 80521, USA\\\textsuperscript{3} Institute for Environment and Society, Brown University, Providence, RI 02912, USA\\\textsuperscript{4} Biology Department, Pennsylvania State University, State College, PA 16802, USA}
\keywords{drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records, Heliophia}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\usepackage{lineno}

\linenumbers
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

\authornote{

Correspondence concerning this article should be addressed to J. Grey
Monroe, 307 University Ave, Fort Collins, CO 80521. E-mail:
\href{mailto:monroejg@colostate.edu}{\nolinkurl{monroejg@colostate.edu}}}

\abstract{
\begin{center}\textbf{Summary}\end{center} \par

Explaining variation in life history strategies is an enduring goal of
evolutionary biology and ecology. For plants, annual and perennial life
histories have long been thought to reflect adaptation to environments
that experience alternative drought regimes. Neverthless, empirical
support for this hypothesis from phylogeographic analyses remains
lacking.

\par

Here, we tested classic life history theory in \textit{Heliophila}
(Brassicaceae), a diverse genus of flowering plants native to Africa,
with a phylogentically constrained analysis integrating 34 years of
satellite-based drought detection with 2192 herbaria occurrence records.

\par

We found that annual and perennial life histories have evolved multiple
times in \textit{Heliophila} and that compared to annuals, perennial
species occur in environments where droughts are significantly less
frequent. We also report evidence that annual plants adapt to
predictable drought regimes by escaping drought prone seasons as seeds.

\par

These results provide compelling support for longstanding theoretical
predictions and a more nuanced understanding of life history evolution.
More broadly, this work outlines a scalable framework integrating
herbaria records and remote sensing to address outstanding questions in
evolutionary ecology. 
}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Understanding the causes and consequences of life history variation is a
longstanding goal of ecology and evolutionary biology (Cole, 1954). In
plants, life histories are especially diverse, with some species
completing their life cycle in a number of weeks to others that live for
thousands of years (Brown, 1996). Along this continuum in angiosperms an
important division exists between annuals which complete their seed to
seed life cycle within a single calendar year from perennials which can
persist over multiple years. Annual plants flower once, set seed,
senesce, and then die, spending at least some portion of the year as a
seed, where they are relatively protected from environmental stress. In
contrast, perennial plants can continue vegetative growth after
reproduction and must survive conditions experienced during all seasons
but can also benefit from multiple bouts of reproduction and competative
advantages. These represent fundamentally different life history
strategies and predicting their occurance is important for community,
ecosystem, and agricultural ecology. However, the environmental factors
that explain their evolution and distributions remain empirically
uresolved (Friedman \& Rubin, 2015).

Classical theory predicts shorter life spans in environments where adult
mortality is high (Charnov \& Schaffer, 1973; Stearns, 1992; Franco \&
Silvertown, 1996). Because lack of water is perhaps the greatest threat
to survival during vegetative or reproductive growth in plants, this
theory has been extended to the hypothesis that annuality is adaptive
when it allows plants to escape drought (Schaffer \& Gadgil, 1975).
Indeed, adaptation to drought, defined as episodes of increased aridity
causing plant sresss (Passioura, 1996), is often invoked as an
explanation for the success of annual species. And while a few cases are
cited where annuality appears to be more common in environments with
greater aridity (Stebbins Jr, 1952; Morishima \emph{et al.}, 1984), this
hypothesis has yet to be supported in a phylogenetic context. In one
previous study where this question was addressed phylogenetically,
(Evans \emph{et al.}, 2005) annuals were not found to be associated with
environments that experience more drought. This could be explained by
the relatively small number of species studied and the reliance on a
limited number of weather stations to characterize environments,
highlighting the need to develop more scalable methods to study trait
evolution phylogeographically.

It is also critical to consider another dimension of drought adaptation:
the expectation that annuality is most adaptive when droughts are not
only frequent but also predictable. That is, when the frequency of
drought is particularly high during certain seasons. Such predictability
is important for selection to lead to escape from those seasons which
are paricuarlary drought prone. While there has been at least one
example of annuality associated with environments qualitatively
classified as \enquote{predictable} in a general sense (Datson \emph{et
al.}, 2008), the seasonal predictability of drought experienced by
annuals has yet to be rigorously studied. As such, further empirical
work is needed to support the model of annuality as a mechanism of
drought adaptation via escape from drought prone seasons.

In addition to drought escape in annuals as a mechanism of adaptation to
frequent predictable droughts, droughts may be necessary for the success
of annuals more generally by acting as episodes of disturbance that
provide opportunites for annuals to establish and compete with sympatric
perennial species. Indeed, there is considerable evidence that
perennials dominate in envirnoments where distrurbance events are
infrequent (Rees \& Long, 1992; Corbin \& D'Antonio, 2004; Clary, 2012).
The resulting prediction from this hypothesis is that in the absence of
frequent drought, perenniality should evolve. However, little is known
about this component of life history evolution because previous work has
largely focused on the origins of annuality rather than perenniality
(Friedman \& Rubin, 2015).

Here we combine a long-term global dataset of satellite detected drought
events with metadata from natural history collections to test these
classic hypotheses within the African endemic mustard genus,
\emph{Heliophila} L. (Brassicaceae). If annuality is an adaptive
strategy allowing plants to escape drought prone seasons, then drought
frequency should predict the distribution of life history strategies
across landscapes, and annual species should be more commonly associated
with drought prone regions than perennial species. Additionally, if
perenniality offers competative advantage in the absence of drought,
associations between life history and drought frequency should be
signficant in even if phylogenies are enriched for transitions from
annual to perenial life history strategy. Finally, if annual species
have adapted to escape predictably drought prone seasons, observations
of growing annual species (i.e.~occurring in forms other than seed)
should be rare during seasons when drought frequency is hightest.
Phylogenetic relatedness can influence tests of associations between
species' traits and their environments (Felsenstein, 1985; Barrett
\emph{et al.}, 1996). Therefore we assessed the relationship between
life history distribution and drought frequency in a phylogenetic
context.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{data}{%
\subsection{Data}\label{data}}

\hypertarget{data-availability}{%
\subsubsection{Data availability}\label{data-availability}}

All analyses were performed using R. All data and the source code to
produce this manuscript are available at
\url{https://github.com/greymonroe/heliophila}. Software used is listed
in the supplement.

\hypertarget{life-history-data-for-heliophila}{%
\subsubsection{\texorpdfstring{Life history data for
\emph{Heliophila}}{Life history data for Heliophila}}\label{life-history-data-for-heliophila}}

\emph{Heliophila} is a genus of flowering plants endemic to the southern
Africa including the Cape Floristic and Succulent Karoo Regions. These
are among the most botanically diverse environments on Earth and the
\emph{Heliophila} species occurring there are considered to make up the
most diverse genus of the family Brassicaceae (Mummenhoff \emph{et al.},
2005; Mandáková \emph{et al.}, 2012). This genus includes both perennial
(n =21) and annual (n = 21) species and this change in life history
strategy has likely arisen multiple independent times (Appel \&
Al-Shehbaz, 1997; Mummenhoff \emph{et al.}, 2005). Furthermore, the fine
scale climatic heterogeneity of Southern Africa is ideal for studying
the distribution of traits in relation to environmental parameters
(Sayre \emph{et al.}, 2013). We used life histories reported by
Mummenhoff \emph{et al.} (2005), grouping species with annual or
perennial life histories. Perenniality was defined based any form of
perennial life history (e.g., herbs, shrubs, mixed, etc). Because the
nature of species reported with mixed traits were unknown
(i.e.~plasticity vs.~genetic variation), we classified these species
here as perennial since they can maintain vegetative growth after
reproduction at least to some capacity.

\hypertarget{heliophila-occurrence-records}{%
\subsubsection{\texorpdfstring{\emph{Heliophila} occurrence
records}{Heliophila occurrence records}}\label{heliophila-occurrence-records}}

To characterize the distributions of annual and perennial
\emph{Heliophila} species, all (8670) records for the genus
\emph{Heliophila} were downloaded from the Global Biodiversity
Information Facility (gbif.org) on July 21, 2018 (GBIF, 2018). herbaria
records records such as these provide a rich data sources to
characterize the geographical distributions of species (Thiers, 2016;
Willis \emph{et al.}, 2017; Lang \emph{et al.}, 2018). And as they
become digitized (Soltis, 2017), these collections have been used to
study relationships between trait distributions, geography, and climate
(Davis \emph{et al.}, 2015; Stropp \emph{et al.}, 2016; Wolf \emph{et
al.}, 2016; Václavı'k \emph{et al.}, 2017).

\hypertarget{sequence-data-for-phylogeny}{%
\subsubsection{Sequence data for
phylogeny}\label{sequence-data-for-phylogeny}}

An alignment of ITS I and II sequences for 21 annual and 21 perennial
\emph{Heliophila} species was obtained from the authors of Mandáková
\emph{et al.} (2012). Individual ITS I and II sequences for
\emph{Aethionema grandiflorum, Alliaria petiolata, Cardamine matthioli,
Chamira circaeoides}, and \emph{Rorippa amphibia} were downloaded from
Genbank.

\hypertarget{satellite-detected-drought-data}{%
\subsubsection{Satellite-detected drought
data}\label{satellite-detected-drought-data}}

Remotely sensed data is a powerful tool for characterizing seasonal
patterns in drought because it is less limited in spatial and temporal
scope and resolution than weather stations or field observations
(AghaKouchak \emph{et al.}, 2015). From an ecological perspective,
droughts are best defined as episodes of plant stress caused by elevated
aridity (Passioura, 1996). Thus remote sensing offers the additional
benefit for studying drought as an agent of natural selection because
plant stress caused by drought can be observed from space (Kogan,
1995a). The remotely sensed Vegetative Health Index (VHI) is one such
metric, which detects landscape scale reductions in plant cover and
temperature conditions characteristic of drought (Kogan, 2001).
Generated from data collected by NOAA AVHRR satellites since 1981, the
VHI is a composite index combining Normalized Difference Vegetation
Index (NDVI) derived quantification of vegetative stress (Vegetative
Condition Index - VCI) with temperature stress indicated by anomalies in
thermal spectra (Temperature Condition Index - TCI). These indicies were
developed to create an unbiased quantification of drought across
ecosystem types. The VHI of year \(y\) during week \(w\) of \([1,52]\)
at pixel \(i\) is derived from the following equations, where \(n\) is
the number of years observed.

\[VCI_{y,w,i} = 100\frac{NDVI_{y,w,i} - NDVI_{min,w,i}}{NDVI_{max,w,i} - NDVI_{min,w,i}}\]

Low values of VCI indicate episodes when plant cover is particularly low
for a given location during a given time of the year. Thus, it controls
for the location and season in quantifying plant stress.

\[TCI_{y,w,i} = 100\frac{T_{max,w,i}-T_{y,w,i}}{T_{max,w,i} - T_{min,w,i}}\]
Similarly, low TCI values indicate episdoes of high thermal stress shown
to be negatively correlated with precipition and soil moisture
(AghaKouchak \emph{et al.}, 2015).

\[VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})\] By combining VCI
and TCI, the VHI is distinguishes drought from other forms of vegetative
stress (Kogan, 1995b). The use of the VHI to detect drought has been
validated globally and across ecosystem types (AghaKouchak \emph{et
al.}, 2015), including in southern Africa, the focal region of this
study (Figure @ref:fig(mapsdroughtexamples)). To date, the VHI has most
often been applied for evaluating drought risk for agricultural research
(e.g., Rojas \emph{et al.}, 2011; Kogan \emph{et al.}, 2016). But it
also presents a tool to study seasonal patterns in the frequency of
drought across environments and to test hypotheses about the effect of
drought on ecological and evolutionary processes (Kerr \& Ostrovsky,
2003). As such, the VHI has been applied recently to study drought
related ecology of natural species and proven useful for predicting
intraspecific variation in drought tolerance traits and genes (Mojica
\emph{et al.}, 2016; Dittberner \emph{et al.}, 2018; Monroe \emph{et
al.}, 2018b). Here, we accessed VHI data at \(16km^2\) resolution from
1981 to 2015
(\url{https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php}) to
characterize the seasonal drought frequencies experienced by annual and
perennial \emph{Heliophila} species across their native range of
southern Africa.

\hypertarget{analyses}{%
\subsection{Analyses}\label{analyses}}

\hypertarget{drought-frequency-calculations}{%
\subsubsection{Drought frequency
calculations}\label{drought-frequency-calculations}}

To characterize drought regimens across the distributions of annual and
perennial species of \emph{Heliophila}, we calculated drought during
different seasons at the location of observations for \emph{Heliophila}
records using the VHI. Specifically, we created global maps of the
frequencies of observing drought conditions between years
(VHI\textless{}40, NOAA) during the winter (quarter surrounding winter
solstice), spring (quarter surrounding spring equinox), summer (quarter
surrounding summer solstice) and fall (quarter surrounding fall equinox)
from 1981 to 2015. From these maps, the drought frequency during the
winter, spring, summer, and fall were extracted for the locations of all
GBIF records.

\hypertarget{filtering-of-occurrence-records}{%
\subsubsection{Filtering of occurrence
records}\label{filtering-of-occurrence-records}}

To avoid instances with spurious location data, we filtered raw GBIF by
restricting our analyses to include only:

\begin{itemize}
\tightlist
\item
  records for species with reported life history\\
\item
  records with geospatial data\\
\item
  records without known geospatial coordinate issues (i.e., coordinates
  reported are those of herbarium)\\
\item
  records from collection sites classified as land pixels in the VHI
  dataset\\
\item
  records from Africa (to exclude locations of cultivation)
\item
  records without duplicates (i.e., identical species, location,
  collection date)
\end{itemize}

\hypertarget{phylogeny-construction-and-ancestral-state-estimation}{%
\subsubsection{Phylogeny construction and ancestral state
estimation}\label{phylogeny-construction-and-ancestral-state-estimation}}

Outgroup ( \emph{Aethionema grandiflorum, Alliaria petiolata, Cardamine
matthioli, Chamira circaeoides}, and \emph{Rorippa amphibia}) and
ingroup \emph{Heliophila} ITS I and II sequences were aligned using
MAFFT (Katoh \emph{et al.}, 2002) with strategy G-INS-I, offset value
0.1, and all other options set as default. The \(GTR + \Gamma\) model of
nucleotide substitution was determined to best fit the data based on AIC
using jModelTest2 (Guindon \& Gascuel, 2003; Darriba \emph{et al.},
2012). A maximum clade credibility tree with branch lengths as relative
time was estimated by summarizing data from six runs of 100,000,000
generations of Bayesian Markov chain Monte Carlo conducted in BEAST 2
(Bouckaert \emph{et al.}, 2014). Model selection and phylogenetic
analyses were conducted through the CIPRES Science Gateway (Miller
\emph{et al.}, 2010). Ancestral state estimation was performed in R
using the package phytools (Revell, 2012) to genearte 10,000 stochastic
character maps simulated under an equal rates model of character
evolution for the trait life habit (annual or perennial).

\hypertarget{comparison-of-drought-frequency-between-annual-and-perennial-species}{%
\subsubsection{Comparison of drought frequency between annual and
perennial
species}\label{comparison-of-drought-frequency-between-annual-and-perennial-species}}

To evaluate the hypothesis that annual and perennial life history
strategies reflect adaptations to alternative drought regimes, we tested
the corresponding prediction that the observed distributions of annual
and perennial \emph{Heliophila} species would be significantly
associated with historic drought frequency. First, we compared the
frequency of drought across all occurrence records of between annual and
perennial species by t-tests. Independent tests were performed to
compare drought frequency between annuals and perennials during each
season. We also tested for a relationship between drought frequency and
life history, season, and their interaction by an analysis of variance
while including species as a random effect. To account for variation in
the number of occurrence records per species, we next calculated the
mean drought frequency during the winter, spring, summer and fall for
each species. Because shared evolutionary history of closely related
species can lead to spurious associations between traits and
environments (Felsenstein, 1985), we tested for a relationship between
life history strategy and drought frequency while controlling for
phylogeny using phylogenetic logistic regression (Ives \& Garland,
2010). This statistical approach is designed to control for the
confounding effects of geographic relationships explained by demographic
history rather than processes of natural selection.

\hypertarget{collection-dates}{%
\subsubsection{Collection dates}\label{collection-dates}}

To test the hypothesis that annual species have adapted to escape
drought prone seasons as seeds, collection dates for herbarium specimens
were compared between annual and perennial species. Comparisons of
distributions were made by Two-sample Kolmogorov-Smirnov test and
Barlett variance test.

\hypertarget{results}{%
\section{Results}\label{results}}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/phylogeny} \caption{Species and examples of herbaria specimens of
\emph{Heliophila} (a) Phylogeny and life history strategies of species
studied. Orange circles at branch tips mark annual species and blue
circles mark perennial species. At internal nodes, pie charts indicate
the estimated posterior probability of being annual versus perennial.
Example herbaria specimens accessed via GBIF of (a) \emph{H. minima},
(b) \emph{H. deserticola}, (c) \emph{H. coronopifolia} and (d) \emph{H.
ephemera}. Images (a,c,d) courtesy of The Bavarian Natural History
Collections (CC BY-SA 4.0) and (b) The London Natural History Museum (CC
BY 4.0). Links to images are found in the supplement.}\label{fig:phylogeny}
\end{figure}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/maps_boxplots} \caption{Locations of occurrence records of (a) annual and (b)
perennial \emph{Heliophila.} Drought frequency during the (c) winter,
(d) spring, (e) summer and (f) fall detected using the VHI. (g) Drought
frequencies during each season at the observation locations of annual
and perennial \emph{Heliophila} (t tests contrasting annuals and
perennials, ** = p \textless{} 0.01).}\label{fig:mapsboxplots}
\end{figure}

The \emph{Heliophila} phylogeny was consistent with previous studies
(Mummenhoff \emph{et al.}, 2005; Mandáková \emph{et al.}, 2012). Based
on 10,000 stochastic character maps simulated under an equal rates model
of character evolution in life history, an average of approximately
eight changes from annual to perennial and five changes from perennial
to annual are observed per stochastic character map (Figure
\ref{fig:phylogeny}a). These results suggest that the ancestral state of
\emph{Heliophila} was annual and that both charcter states have arisen
independently mutliple times.

Out of 8670 \emph{Heliophila} GBIF records, 6634 were for species with
reported life history (Mummenhoff \emph{et al.}, 2005), 2856 had
geospatial data, 2833 did not have geospatial issues, 2684 were located
on pixels classified as land having drought measurements, 2543 were
located in Africa, 2192 were not duplicated. Thus, after all filtering
steps, 2192 records for 42 species (Figure \ref{fig:phylogeny}, Table
\ref{tab:speciesmeanstable}) passed for further analyses. The number of
samples varied between species, with a mean of 52.19 samples per
species. \emph{H. rigidiuscula} had the most records, 201, and \emph{H.
cornellsbergia} the fewest, 2 (Table \ref{tab:speciesmeanstable}).

There were clear visual differences between the distributions of the 960
annual and the 1232 perennial \emph{Heliophila} observation records (see
Figure \ref{fig:speciesmaps} for maps of individual species). While
annual species were generally found in the western regions of South
Africa and Namibia, primarily in the Cape Floristic Region and Succulent
Karoo (Figure \ref{fig:mapsboxplots}a), the occurrence of perennials
extended to the southern and eastern coast of South Africa (Figure
\ref{fig:mapsboxplots}b).

The frequency of drought varied considerably across the ranges of
\emph{Heliophila} species (Figure \ref{fig:mapsboxplots}c-f). This
heterogeneity is expected, given that this is one of the most
climatically diverse regions of the Earth (Sayre \emph{et al.}, 2013).
It is worth noting the east to west cline in drought frequency observed
during the summer, which distinguishes the high drought frequency of the
Kalahari Sands and Namid Desert phytogeographic regions from the low
drought frequency of the Drakensberg Mountains and Coastal Zambesian
phytogeographic regions. In the Cape phytogeographic region there was
finer scale heterogeneity in drought frequency during the summer.

We found that the frequency of drought was significantly higher at the
locations of occurrence records for annual species. When comparing
across all occurrence records (all records rather than species means,
Figure \ref{fig:mapsboxplots}g), the frequency of drought was
significantly higher at the location of annuals during the winter (t =
10.65, p \textless{} 0.01), spring (t = 10.73, p \textless{} 0.01),
summer (t = 12.67, p = \textless{} 0.01), and fall (t = 15.26, p
\textless{} 0.01). A mixed-model analysis of variace which included
species as random effect also identified a signficant relationship
between drought frequency and life history (p \textless{} 0.01), season
(p \textless{} 0.01) and their interaction (p \textless{} 0.01) (Table
\ref{tab:anovatable}). Because a comparison across all occurrence
records does not account for variation in the number of records per
species (Table \ref{tab:speciesmeanstable}) or species relatedness
(Figure \ref{fig:phylogeny}a), we also tested whether mean drought
frequency values of each species were significantly different between
annuals and perennials using phylogenetic logistic regression. We found
that the mean drought frequencies were significantly higher
(\(\alpha = 0.05\)) in annual species during the spring, summer, and
fall (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a,b). These
findings indicate that demographic history alone does not explain
differences in the drought frequencies experienced between the
environments of annual and perennial \emph{Heliophila}.

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:modelstable}Phylogenetic logistic regressions between life history, and the mean drought frequency observed at specimen sites of Heliophila species the winter, spring, summer, and fall.}
\begin{tabular}{lll}
\toprule
Predictor & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{P}\\
\midrule
Intercept & 0.7231 & 0.6636\\
Winter drought freq. & -1.5452 & 0.7274\\ \midrule
Intercept & 5.0107 & 0.0534\\
Spring drought freq. & -12.9014 & 0.0464\\ \midrule
Intercept & 7.7093 & 0.0054\\
Summer drought freq. & -19.9056 & 0.0042\\ \midrule
Intercept & 7.0162 & 0.0082\\
Fall drought freq. & -20.8174 & 0.0067\\ \midrule
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Annual species were scored as 0 and perennial species as 1.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/line_and_dates} \caption{(a) Comparison (mean +- SE) of drought frequency across
seasons calcualted at the occurance locations of GBIF records of annual
and perennial species of \emph{Heliophila}. (b) Results from
phylogenetic logistic regression, where the line shows the model fit and
* = p \textless{} 0.05, ** = p \textless{} 0.01. Annuals were scored as
0 and perennials as 1. (c) Collection dates of GBIF records of annual
and perennial species of \emph{Heliophila} in relation to the
photoperiodic calender where day 1 is intermediate to the fall equinox
and winter soltice.}\label{fig:lineplots}
\end{figure}

The preceding results indicate that annual species are found in
environments where droughts are significantly more frequent, especially
in the summer and fall. Classic life history theory hypothesizes that
annuality reflects adaptation to such environments because it allows
species to escape predictable stressful conditions. If this is the case,
we would expect that annuals spend the drought prone seasons of summer
and fall as seeds. To test this hypothesis, we compared the dates of
occurrence records between annual and perennial \emph{Heliophila}
species. The distributions reveal a considerable difference in the
timing of observation of these two life histories. In comparison to
perennials, which appear to be collected throughout the year, annuals
are almost exclusively observed during the winter and spring (Figure
\ref{fig:lineplots}b). The differences between the distribution of
collection dates were significant by all tests (ks.test D = 0.25, p
\textless{} 0.01; bartlett.test K2 = 503.18, p \textless{} 0.01) This is
consistent with a model of life history in which annual species flower
in the spring, set seed, senesce, and die before the summer. Thus, these
annual species are likely to remain dormant during the summer and fall,
when drought is the strongest predictor of the distributions of annual
and perennial life histories (Figure \ref{fig:lineplots}a).

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

To test the hypothesis that annual and perennial plants reflect
adaptation to alternative drought environments we examined the landscape
distribution of life history strategies in the large and diverse mustard
genus, \emph{Heliophila}. Using metadata of 2192 occurrence records and
a 34 year dataset of satellite-detected droughts, we tested the
prediction that annual species are more often observed in drought-prone
locations than perennial species, when controlling for phylogenetic
relatedness. We found that drought frequency is significantly different
between the distributions of annual and perennial species, with annuals
being found in environments with more frequent drought, and that this
signal is strongest during the seasons when annuals are likely escaping
via seed dormancy. These results remain significant while controlling
for the phylogenetic relationships of \emph{Heliophila} species,
yielding support for the role that natural selection has played in
driving contemporary distributions of these alternatives strategies in
relation to drought regimens.

We cannot eliminate the possibility that confounding traits or
environmental variables are the causative factors explaining variation
in the distributions of annual and perennial species. Nevertheless,
these results provide quantitative support for the classic prediction
that annual species are found in environments that experience more
frequent drought than perennial species. These findings are a valuable
contribution to previous investigations of associations between life
history and climate (Morishima \emph{et al.}, 1984; Evans \emph{et al.},
2005; Datson \emph{et al.}, 2008; Cruz-Mazo \emph{et al.}, 2009). To our
knowledge this is the first study to demonstrate a significant
association between life history and drought in a phylogenetic context
informed by large scale species distribution data and long term drought
detection.

Unfortunately, herbarium collections and their associated data do not
represent systematic or random sampling of a species distribution.
Significant biases in collecting exist, which we have not necessarily
controlled for here, and may have some effect on our findings, such as a
bias toward collecting near roads or near the locations of natural
history collections (Daru \emph{et al.}, 2018). Future research will
benefit from systematic sampling efforts to avoid these noted biases.
However, the ecosystems of southern Africa include several biodiversity
hotspots and are among the most botanically well sampled regions on
Earth (Daru \emph{et al.}, 2018), suggesting that this may currently be
the optimal region for our analyses of life history distribution.
Indeed, we were able to use 2192 occurrence records to study 42 species,
which represents a significant advance over relying on personal
observations to characterize species distributions.

These findings empirical provide support classical theoretical
predictions about the adaptive value of annual and perennial life
history strategies. Taken together, they suggest that in
\emph{Heliophila}, annual species are adapted to environments with
increased summer droughts by avoiding these seasons in a dormant seed
phase of their life cycle. They also suggest that perenniality is
adaptive in environemnts where droughts are less frequent. While most
previous work has focused on describing the evolutionary origins of
annuality (Barrett \emph{et al.}, 1996; Conti \emph{et al.}, 1999;
Andreasen \& Baldwin, 2001; Verboom \emph{et al.}, 2004; Friedman \&
Rubin, 2015) there are at least a few other cases where perenniality
appears to have arisen from an annual ancestor (Bena \emph{et al.},
1998; Tank \& Olmstead, 2008). And while early theory predicted
selection for annuality when adult mortality is high (Stearns, 1992), we
also find evidence that perenniality could be explained by reduced
frequency of drought. This is supported by the theoretical prediction
that perenniality is advantageous is stable habitats. The phylogeny
reveals several transitions from annual to perennial life history
(Figure \ref{fig:phylogeny}a) and the distributions of perennial
\emph{Heliophila} extend into regions where drought frequency is low
(Figure \ref{fig:mapsboxplots}b, Figure \ref{fig:speciesmaps}).
Perennials may be able to out compete annual relatives in environments
where the infrequency of drought favors strategies that allow plants to
benefit from growth over many seasons. This also suggests that annuals
rely on drought as a source of disturbance for seedling recruitment when
competing with perennials (Corbin \& D'Antonio, 2004). Indeed, no annual
species were observed in the low drought regions of eastern South Africa
(Figure \ref{fig:mapsboxplots}, Figure \ref{fig:speciesmaps}).

These findings suggest that species with locally adaptive life history
strategies could be threatened by rapidly changing drought regimens
(Dai, 2011). In light of the findings here, changes in eastern South
Africa which is forecast to experience signicantly lower rainfall
(Service \& Comission, 2017) could be particularly impactful. In this
study we found that this region is currently dominiated by derived
perennial species of \emph{Heliophila}. However, if drought become more
frequent in this region, it may allow for the establishement of annuals.
Such changes in selection patterns and shifts in plant functional
diversity could have impacts on ecosystem functioning and processes such
as carbon cycling (Garnier \emph{et al.}, 1997; Roumet \emph{et al.},
2006; Monroe \emph{et al.}, 2018a). Furthermore, the changes in
frequency of drought may be an important factor when considering the use
of perennial cropping systems (Parry \emph{et al.}, 2005; Lelièvre \&
Volaire, 2009).

In conclusion, we find strong support for classic life history theory
which predicts that annuality is adaptive in environments where droughts
occur more frequently. Additionally, we report evidence consistent with
a life history model in annuals in which they escape drought prone
seasons during the seed phase of their life cycle. Finally, we find
evidence that the distributions of perennial lineages may indicate a
competitive advantage in areas where droughts are infrequent. More
broadly, this work highlights the irreplaceable value of natural history
collections and demonstrates the power of combining such information
with large scale remote sensing data to address outstanding classic
hypotheses in ecology and evolution.

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

We thank Jesse Lasky, members of the Sloan lab at CSU and three
ananymous reviewers for generous feedback that improved the quality of
this work. This research was supported by NSF Award 1701918 and
USDA-NIFA Award 2014-38420-21801 to JGM.

\hypertarget{author-contributions}{%
\section{Author contributions}\label{author-contributions}}

JGM, BG, KGT and JKM contributed to the design of the research,
interpretation, and writing the manuscript. JGM, BG, and KGT contributed
to the performance of the research and data analysis.

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-aghakouchak2015remote}{}%
\textbf{\textnormal{AghaKouchak A}, \textnormal{Farahmand A},
\textnormal{Melton F}, \textnormal{Teixeira J}, \textnormal{Anderson M},
\textnormal{Wardlow BD}, \textnormal{Hain C}}. \textbf{2015}. Remote
sensing of drought: Progress, challenges and opportunities.
\emph{Reviews of Geophysics} \textbf{53}: 452--480.

\leavevmode\hypertarget{ref-andreasen2001unequal}{}%
\textbf{\textnormal{Andreasen K}, \textnormal{Baldwin BG}}.
\textbf{2001}. Unequal evolutionary rates between annual and perennial
lineages of checker mallows (sidalcea, malvaceae): Evidence from
18S--26S rDNA internal and external transcribed spacers. \emph{Molecular
Biology and Evolution} \textbf{18}: 936--944.

\leavevmode\hypertarget{ref-appel1997generic}{}%
\textbf{\textnormal{Appel O}, \textnormal{Al-Shehbaz IA}}.
\textbf{1997}. Generic limits and taxonomy of hornungia, pritzelago, and
hymenolobus (brassicaceae). \emph{Novon}: 338--340.

\leavevmode\hypertarget{ref-barrett1996comparative}{}%
\textbf{\textnormal{Barrett SCH}, \textnormal{Harder LD},
\textnormal{Worley AC}}. \textbf{1996}. The comparative biology of
pollination and mating in flowering plants. \emph{Phil. Trans. R. Soc.
Lond. B} \textbf{351}: 1271--1280.

\leavevmode\hypertarget{ref-bena1998molecular}{}%
\textbf{\textnormal{Bena G}, \textnormal{Lejeune B},
\textnormal{Prosperi J-M}, \textnormal{Olivieri I}}. \textbf{1998}.
Molecular phylogenetic approach for studying life-history evolution: The
ambiguous example of the genus medicago l. \emph{Proceedings of the
Royal Society of London B: Biological Sciences} \textbf{265}:
1141--1151.

\leavevmode\hypertarget{ref-bouckaert2014beast}{}%
\textbf{\textnormal{Bouckaert R}, \textnormal{Heled J},
\textnormal{Kühnert D}, \textnormal{Vaughan T}, \textnormal{Wu C-H},
\textnormal{Xie D}, \textnormal{Suchard M}, \textnormal{Rambaut A},
\textnormal{Drummond A}}. \textbf{2014}. BEAST 2: A software platform
for bayesian evolutionary analysis. \emph{PLoS Computational Biology}
\textbf{10}: doi:10.1371/journal.pcbi.1003537.

\leavevmode\hypertarget{ref-brown1996oldlist}{}%
\textbf{\textnormal{Brown PM}}. \textbf{1996}. OLDLIST: A database of
maximum tree ages. \emph{Tree rings, environment, and humanity.
Radiocarbon} \textbf{1996}: 727--731.

\leavevmode\hypertarget{ref-charnov1973life}{}%
\textbf{\textnormal{Charnov EL}, \textnormal{Schaffer WM}}.
\textbf{1973}. Life-history consequences of natural selection: Cole's
result revisited. \emph{The American Naturalist} \textbf{107}: 791--793.

\leavevmode\hypertarget{ref-clary2012determinants}{}%
\textbf{\textnormal{Clary J}}. \textbf{2012}. Determinants of perennial
and annual grass distribution in mediterranean-climate california.
\emph{Plant Ecology} \textbf{213}: 1203--1208.

\leavevmode\hypertarget{ref-cole1954population}{}%
\textbf{\textnormal{Cole LC}}. \textbf{1954}. The population
consequences of life history phenomena. \emph{The Quarterly Review of
Biology} \textbf{29}: 103--137.

\leavevmode\hypertarget{ref-conti1999phylogenetic}{}%
\textbf{\textnormal{Conti E}, \textnormal{Soltis DE}, \textnormal{Hardig
TM}, \textnormal{Schneider J}}. \textbf{1999}. Phylogenetic
relationships of the silver saxifrages (saxifraga, sect. Ligulatae
haworth): Implications for the evolution of substrate specificity, life
histories, and biogeography. \emph{Molecular Phylogenetics and
Evolution} \textbf{13}: 536--555.

\leavevmode\hypertarget{ref-corbin2004competition}{}%
\textbf{\textnormal{Corbin JD}, \textnormal{D'Antonio CM}}.
\textbf{2004}. Competition between native perennial and exotic annual
grasses: Implications for an historical invasion. \emph{Ecology}
\textbf{85}: 1273--1283.

\leavevmode\hypertarget{ref-cruz2009molecular}{}%
\textbf{\textnormal{Cruz-Mazo G}, \textnormal{Buide M},
\textnormal{Samuel R}, \textnormal{Narbona E}}. \textbf{2009}. Molecular
phylogeny of scorzoneroides (asteraceae): Evolution of heterocarpy and
annual habit in unpredictable environments. \emph{Molecular
phylogenetics and evolution} \textbf{53}: 835--847.

\leavevmode\hypertarget{ref-dai2011drought}{}%
\textbf{\textnormal{Dai A}}. \textbf{2011}. Drought under global
warming: A review. \emph{Wiley Interdisciplinary Reviews: Climate
Change} \textbf{2}: 45--65.

\leavevmode\hypertarget{ref-darriba2012jmodeltest}{}%
\textbf{\textnormal{Darriba D}, \textnormal{Taboada G},
\textnormal{Doallo R}, \textnormal{Posada D}}. \textbf{2012}. JModelTest
2: More models, new heuristics and parallel computing. \emph{Nature
Methods} \textbf{9}: 772.

\leavevmode\hypertarget{ref-daru2018widespread}{}%
\textbf{\textnormal{Daru BH}, \textnormal{Park DS}, \textnormal{Primack
RB}, \textnormal{Willis CG}, \textnormal{Barrington DS},
\textnormal{Whitfeld TJ}, \textnormal{Seidler TG}, \textnormal{Sweeney
PW}, \textnormal{Foster DR}, \textnormal{Ellison AM} \emph{et al.}}
\textbf{2018}. Widespread sampling biases in herbaria revealed from
large-scale digitization. \emph{New Phytologist} \textbf{217}: 939--955.

\leavevmode\hypertarget{ref-datson2008climate}{}%
\textbf{\textnormal{Datson P}, \textnormal{Murray B},
\textnormal{Steiner K}}. \textbf{2008}. Climate and the evolution of
annual/perennial life-histories in nemesia (scrophulariaceae).
\emph{Plant Systematics and Evolution} \textbf{270}: 39--57.

\leavevmode\hypertarget{ref-davis2015herbarium}{}%
\textbf{\textnormal{Davis CC}, \textnormal{Willis CG},
\textnormal{Connolly B}, \textnormal{Kelly C}, \textnormal{Ellison AM}}.
\textbf{2015}. Herbarium records are reliable sources of phenological
change driven by climate and provide novel insights into species'
phenological cueing mechanisms. \emph{American Journal of Botany}
\textbf{102}: 1599--1609.

\leavevmode\hypertarget{ref-dittberner2018natural}{}%
\textbf{\textnormal{Dittberner H}, \textnormal{Korte A},
\textnormal{Mettler-Altmann T}, \textnormal{Weber A}, \textnormal{Monroe
G}, \textnormal{Meaux J de}}. \textbf{2018}. Natural variation in
stomata size contributes to the local adaptation of water-use efficiency
in arabidopsis thaliana. \emph{bioRxiv}: 253021.

\leavevmode\hypertarget{ref-evans2005climate}{}%
\textbf{\textnormal{Evans ME}, \textnormal{Hearn DJ}, \textnormal{Hahn
WJ}, \textnormal{Spangle JM}, \textnormal{Venable DL}}. \textbf{2005}.
Climate and life-history evolution in evening primroses (oenothera,
onagraceae): A phylogenetic comparative analysis. \emph{Evolution}
\textbf{59}: 1914--1927.

\leavevmode\hypertarget{ref-felsenstein1985phylogenies}{}%
\textbf{\textnormal{Felsenstein J}}. \textbf{1985}. Phylogenies and the
comparative method. \emph{American Naturalist} \textbf{125}: 1--15.

\leavevmode\hypertarget{ref-franco1996life}{}%
\textbf{\textnormal{Franco M}, \textnormal{Silvertown J}}.
\textbf{1996}. Life history variation in plants: An exploration of the
fast-slow continuum hypothesis. \emph{Phil. Trans. R. Soc. Lond. B}
\textbf{351}: 1341--1348.

\leavevmode\hypertarget{ref-friedman2015all}{}%
\textbf{\textnormal{Friedman J}, \textnormal{Rubin MJ}}. \textbf{2015}.
All in good time: Understanding annual and perennial strategies in
plants. \emph{American journal of botany} \textbf{102}: 497--499.

\leavevmode\hypertarget{ref-garnier1997specific}{}%
\textbf{\textnormal{Garnier E}, \textnormal{Cordonnier P},
\textnormal{Guillerm J-L}, \textnormal{Sonié L}}. \textbf{1997}.
Specific leaf area and leaf nitrogen concentration in annual and
perennial grass species growing in mediterranean old-fields.
\emph{Oecologia} \textbf{111}: 490--498.

\leavevmode\hypertarget{ref-gbifdownload}{}%
\textbf{\textnormal{GBIF}}. \textbf{2018}. GBIF occurrence download.

\leavevmode\hypertarget{ref-guindon2003a}{}%
\textbf{\textnormal{Guindon S}, \textnormal{Gascuel O}}. \textbf{2003}.
A simple, fast and accurate method to estimate large phylogenies by
maximum-likelihood. \emph{Systematic Biology} \textbf{52}: 696--704.

\leavevmode\hypertarget{ref-ives2010phylogenetic}{}%
\textbf{\textnormal{Ives A}, \textnormal{Garland T}}. \textbf{2010}.
CPhylogenetic logistic regression for binary dependent variables.
\emph{Systematic Biology} \textbf{59}: 9--26.

\leavevmode\hypertarget{ref-katoh2002mafft}{}%
\textbf{\textnormal{Katoh}, \textnormal{Misawa}, \textnormal{Kuma},
\textnormal{Miyata}}. \textbf{2002}. MAFFT: A novel method for rapid
multiple sequence alignment based on fast fourier transform.
\emph{Nucleic Acids Research} \textbf{30}: 3059--3066.

\leavevmode\hypertarget{ref-kerr2003space}{}%
\textbf{\textnormal{Kerr JT}, \textnormal{Ostrovsky M}}. \textbf{2003}.
From space to species: Ecological applications for remote sensing.
\emph{Trends in ecology \& evolution} \textbf{18}: 299--305.

\leavevmode\hypertarget{ref-kogan1995droughts}{}%
\textbf{\textnormal{Kogan FN}}. \textbf{1995a}. Droughts of the late
1980s in the united states as derived from noaa polar-orbiting satellite
data. \emph{Bulletin of the American Meteorological Society}
\textbf{76}: 655--668.

\leavevmode\hypertarget{ref-kogan1995application}{}%
\textbf{\textnormal{Kogan FN}}. \textbf{1995b}. Application of
vegetation index and brightness temperature for drought detection.
\emph{Advances in space research} \textbf{15}: 91--100.

\leavevmode\hypertarget{ref-kogan2001operational}{}%
\textbf{\textnormal{Kogan FN}}. \textbf{2001}. Operational space
technology for global vegetation assessment. \emph{Bulletin of the
American Meteorological Society} \textbf{82}: 1949--1964.

\leavevmode\hypertarget{ref-kogan2016modelling}{}%
\textbf{\textnormal{Kogan F}, \textnormal{Guo W}, \textnormal{Strashnaia
A}, \textnormal{Kleshenko A}, \textnormal{Chub O}, \textnormal{Virchenko
O}}. \textbf{2016}. Modelling and prediction of crop losses from noaa
polar-orbiting operational satellites. \emph{Geomatics, Natural Hazards
and Risk} \textbf{7}: 886--900.

\leavevmode\hypertarget{ref-lang2018using}{}%
\textbf{\textnormal{Lang PL}, \textnormal{Willems FM},
\textnormal{Scheepens J}, \textnormal{Burbano HA}, \textnormal{Bossdorf
O}}. \textbf{2018}. \emph{Using herbaria to study global environmental
change}. PeerJ Preprints.

\leavevmode\hypertarget{ref-lelievre2009current}{}%
\textbf{\textnormal{Lelièvre F}, \textnormal{Volaire F}}. \textbf{2009}.
Current and potential development of perennial grasses in rainfed
mediterranean farming systems. \emph{Crop Science} \textbf{49}:
2371--2378.

\leavevmode\hypertarget{ref-mandakova2012whole}{}%
\textbf{\textnormal{Mandáková T}, \textnormal{Mummenhoff K},
\textnormal{Al-Shehbaz IA}, \textnormal{Mucina L},
\textnormal{Mühlhausen A}, \textnormal{Lysak MA}}. \textbf{2012}.
Whole-genome triplication and species radiation in the southern african
tribe heliophileae (brassicaceae). \emph{Taxon} \textbf{61}: 989--1000.

\leavevmode\hypertarget{ref-miller2010creating}{}%
\textbf{\textnormal{Miller M}, \textnormal{Pfeiffer W},
\textnormal{Schwartz T and}}. \textbf{2010}. Creating the cipres science
gateway for inference of large phylogenetic trees. \emph{Proceedings of
the Gateway Computing Environments Workshop}: 1--8.

\leavevmode\hypertarget{ref-Mojica2016}{}%
\textbf{\textnormal{Mojica JP}, \textnormal{Mullen J},
\textnormal{Lovell JT}, \textnormal{Monroe JG}, \textnormal{Paul JR},
\textnormal{Oakley CG}, \textnormal{McKay JK}}. \textbf{2016}. Genetics
of water use physiology in locally adapted Arabidopsis thaliana.
\emph{Plant Science}.

\leavevmode\hypertarget{ref-monroe2018ecoevolutionary}{}%
\textbf{\textnormal{Monroe J}, \textnormal{Markman D}, \textnormal{Beck
W}, \textnormal{Felton A}, \textnormal{Vahsen M}, \textnormal{Pressler
Y}}. \textbf{2018a}. Ecoevolutionary dynamics of carbon cycling in the
anthropocene. \emph{Trends in ecology \& evolution} \textbf{33}:
213--225.

\leavevmode\hypertarget{ref-monroe2018drought}{}%
\textbf{\textnormal{Monroe J}, \textnormal{Powell T}, \textnormal{Price
N}, \textnormal{Mullen J}, \textnormal{Howard A}, \textnormal{Evans K},
\textnormal{Lovell J}, \textnormal{McKay J}}. \textbf{2018b}. Drought
adaptation in nature by extensive genetic loss-of-function.
\emph{eLife}: DOI: 10.7554/eLife.41038.

\leavevmode\hypertarget{ref-monyela2017two}{}%
\textbf{\textnormal{Monyela BM}}. \textbf{2017}. A two-year long drought
in summer 2014/2015 and 2015/2016 over south africa.

\leavevmode\hypertarget{ref-morishima1984differentiation}{}%
\textbf{\textnormal{Morishima H}, \textnormal{Sano Y}, \textnormal{Oka
H}}. \textbf{1984}. Differentiation of perennial and annual types due to
habitat conditions in the wild rice oryza perennis. \emph{Plant
Systematics and Evolution} \textbf{144}: 119--135.

\leavevmode\hypertarget{ref-mummenhoff2005phylogeny}{}%
\textbf{\textnormal{Mummenhoff K}, \textnormal{Al-Shehbaz IA},
\textnormal{Bakker FT}, \textnormal{Linder HP}, \textnormal{Mühlhausen
A}}. \textbf{2005}. Phylogeny, morphological evolution, and speciation
of endemic brassicaceae genera in the cape flora of southern africa.
\emph{Annals of the Missouri Botanical Garden}: 400--424.

\leavevmode\hypertarget{ref-parry2005prospects}{}%
\textbf{\textnormal{Parry M}, \textnormal{Flexas J}, \textnormal{Medrano
H}}. \textbf{2005}. Prospects for crop production under drought:
Research priorities and future directions. \emph{Annals of Applied
Biology} \textbf{147}: 211--226.

\leavevmode\hypertarget{ref-passioura1996drought}{}%
\textbf{\textnormal{Passioura J}}. \textbf{1996}. Drought and drought
tolerance. \emph{Plant growth regulation} \textbf{20}: 79--83.

\leavevmode\hypertarget{ref-rees1992germination}{}%
\textbf{\textnormal{Rees M}, \textnormal{Long MJ}}. \textbf{1992}.
Germination biology and the ecology of annual plants. \emph{The American
Naturalist} \textbf{139}: 484--508.

\leavevmode\hypertarget{ref-revell2012phytools}{}%
\textbf{\textnormal{Revell LJ}}. \textbf{2012}. Phytools: An r package
for phylogenetic comparative biology (and other things). \emph{Methods
in Ecology and Evolution} \textbf{3}: 217--223.

\leavevmode\hypertarget{ref-rojas2011assessing}{}%
\textbf{\textnormal{Rojas O}, \textnormal{Vrieling A},
\textnormal{Rembold F}}. \textbf{2011}. Assessing drought probability
for agricultural areas in africa with coarse resolution remote sensing
imagery. \emph{Remote sensing of Environment} \textbf{115}: 343--352.

\leavevmode\hypertarget{ref-roumet2006suites}{}%
\textbf{\textnormal{Roumet C}, \textnormal{Urcelay C}, \textnormal{Dı'az
S}}. \textbf{2006}. Suites of root traits differ between annual and
perennial species growing in the field. \emph{New phytologist}
\textbf{170}: 357--368.

\leavevmode\hypertarget{ref-sayre2013new}{}%
\textbf{\textnormal{Sayre RG}, \textnormal{Comer P}, \textnormal{Hak J},
\textnormal{Josse C}, \textnormal{Bow J}, \textnormal{Warner H},
\textnormal{Larwanou M}, \textnormal{Kelbessa E}, \textnormal{Bekele T},
\textnormal{Kehl H} \emph{et al.}} \textbf{2013}. A new map of
standardized terrestrial ecosystems of africa. \emph{African
Geographical Review}.

\leavevmode\hypertarget{ref-schaffer1975selection}{}%
\textbf{\textnormal{Schaffer W}, \textnormal{Gadgil M}}. \textbf{1975}.
Selection for optimal life histories in plants. \emph{Ecology and
evolution of communities.}: 142--157.

\leavevmode\hypertarget{ref-south2017a}{}%
\textbf{\textnormal{Service SAW}, \textnormal{Comission WR}}.
\textbf{2017}. \emph{A climate change reference atlas}.

\leavevmode\hypertarget{ref-soltis2017digitization}{}%
\textbf{\textnormal{Soltis PS}}. \textbf{2017}. Digitization of herbaria
enables novel research. \emph{American journal of botany} \textbf{104}:
1281--1284.

\leavevmode\hypertarget{ref-stearns1992evolution}{}%
\textbf{\textnormal{Stearns SC}}. \textbf{1992}. \emph{The evolution of
life histories}.

\leavevmode\hypertarget{ref-stebbins1952aridity}{}%
\textbf{\textnormal{Stebbins Jr GL}}. \textbf{1952}. Aridity as a
stimulus to plant evolution. \emph{The American Naturalist} \textbf{86}:
33--44.

\leavevmode\hypertarget{ref-stropp2016mapping}{}%
\textbf{\textnormal{Stropp J}, \textnormal{Ladle RJ}, \textnormal{M.
Malhado AC}, \textnormal{Hortal J}, \textnormal{Gaffuri J},
\textnormal{H. Temperley W}, \textnormal{Olav Skøien J},
\textnormal{Mayaux P}}. \textbf{2016}. Mapping ignorance: 300 years of
collecting flowering plants in africa. \emph{Global Ecology and
Biogeography} \textbf{25}: 1085--1096.

\leavevmode\hypertarget{ref-tank2008annuals}{}%
\textbf{\textnormal{Tank DC}, \textnormal{Olmstead RG}}. \textbf{2008}.
From annuals to perennials: Phylogeny of subtribe castillejinae
(orobanchaceae). \emph{American Journal of Botany} \textbf{95}:
608--625.

\leavevmode\hypertarget{ref-thiers2016index}{}%
\textbf{\textnormal{Thiers B}}. \textbf{2016}. Index herbariorum: A
global directory of public herbaria and associated staff. New york
botanical garden's virtual herbarium. \emph{http://sweetgum. nybg.
org/ih}.

\leavevmode\hypertarget{ref-vaclavik2017effects}{}%
\textbf{\textnormal{Václavı'k T}, \textnormal{Beckmann M},
\textnormal{Cord AF}, \textnormal{Bindewald AM}}. \textbf{2017}. Effects
of uv-b radiation on leaf hair traits of invasive plants---combining
historical herbarium records with novel remote sensing data. \emph{PloS
one} \textbf{12}: e0175671.

\leavevmode\hypertarget{ref-verboom2004testing}{}%
\textbf{\textnormal{Verboom GA}, \textnormal{Linder HP},
\textnormal{Stock WD}}. \textbf{2004}. Testing the adaptive nature of
radiation: Growth form and life history divergence in the african grass
genus ehrharta (poaceae: Ehrhartoideae). \emph{American Journal of
Botany} \textbf{91}: 1364--1370.

\leavevmode\hypertarget{ref-willis2017old}{}%
\textbf{\textnormal{Willis CG}, \textnormal{Ellwood ER},
\textnormal{Primack RB}, \textnormal{Davis CC}, \textnormal{Pearson KD},
\textnormal{Gallinat AS}, \textnormal{Yost JM}, \textnormal{Nelson G},
\textnormal{Mazer SJ}, \textnormal{Rossington NL} \emph{et al.}}
\textbf{2017}. Old plants, new tricks: Phenological research using
herbarium specimens. \emph{Trends in ecology \& evolution} \textbf{32}:
531--546.

\leavevmode\hypertarget{ref-wolf2016altitudinal}{}%
\textbf{\textnormal{Wolf A}, \textnormal{Zimmerman NB},
\textnormal{Anderegg WR}, \textnormal{Busby PE}, \textnormal{Christensen
J}}. \textbf{2016}. Altitudinal shifts of the native and introduced
flora of c alifornia in the context of 20th-century warming.
\emph{Global ecology and biogeography} \textbf{25}: 418--429.

\endgroup

\newpage
\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}

\hypertarget{supplement}{%
\section{Supplement}\label{supplement}}

\hypertarget{images-used}{%
\subsubsection{Images used}\label{images-used}}

\url{https://www.gbif.org/occurrence/1099023487}\\
\url{https://www.gbif.org/occurrence/1057389408}~\\
\url{https://www.gbif.org/occurrence/1099023562}~\\
\url{https://www.gbif.org/occurrence/1099023490}

\hypertarget{supplementary-tables-and-figures}{%
\subsubsection{Supplementary tables and
figures}\label{supplementary-tables-and-figures}}

\newpage

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:anovatable}Analysis of variance (ANOVA) to compare drought frequency as a function of life history, season, and their interaction while including species as a random effect.}
\begin{tabular}{lllll}
\toprule
predictor & \multicolumn{1}{c}{numDF} & \multicolumn{1}{c}{denDF} & \multicolumn{1}{c}{F.value} & \multicolumn{1}{c}{p.value}\\
\midrule
intercept & 1 & 8,720.0000 & 2,833.9567 & 0.0000\\
life history & 1 & 40.0000 & 12.2117 & 0.0012\\
season & 3 & 8,720.0000 & 106.4947 & 0.0000\\
life history x season & 3 & 8,720.0000 & 4.8301 & 0.0023\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\end{table}

\begin{center}
\begin{ThreePartTable}
\begin{TableNotes}[para]
\normalsize{\textit{Note.} LH = Life history (a = annual, p = perennial). n=sample size of GBIF records. Seasons are mean drought frequencies observed at locations of records.}
\end{TableNotes}
\small{
\begin{longtable}{lllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:speciesmeanstable}Heliophila species records and the mean drought frequencies during different seasons at the location of records }\\
\toprule
Species & \multicolumn{1}{c}{LH} & \multicolumn{1}{c}{n} & \multicolumn{1}{c}{Winter} & \multicolumn{1}{c}{Spring} & \multicolumn{1}{c}{Summer} & \multicolumn{1}{c}{Fall}\\
\midrule
Heliophila acuminata & a & 28 & 0.32 & 0.38 & 0.41 & 0.36\\
Heliophila africana & a & 91 & 0.33 & 0.35 & 0.34 & 0.34\\
Heliophila amplexicaulis & a & 60 & 0.32 & 0.36 & 0.39 & 0.33\\
Heliophila arenaria & a & 65 & 0.34 & 0.37 & 0.38 & 0.34\\
Heliophila carnosa & p & 129 & 0.33 & 0.37 & 0.39 & 0.31\\
Heliophila cedarbergensis & p & 3 & 0.40 & 0.43 & 0.32 & 0.27\\
Heliophila collina & a & 16 & 0.35 & 0.47 & 0.48 & 0.45\\
Heliophila cornellsbergia & a & 2 & 0.33 & 0.42 & 0.35 & 0.21\\
Heliophila cornuta & p & 101 & 0.35 & 0.40 & 0.40 & 0.34\\
Heliophila coronopifolia & a & 40 & 0.37 & 0.42 & 0.40 & 0.37\\
Heliophila crithmifolia & a & 97 & 0.35 & 0.42 & 0.45 & 0.38\\
Heliophila descurva & a & 12 & 0.36 & 0.38 & 0.38 & 0.29\\
Heliophila deserticola & a & 133 & 0.48 & 0.48 & 0.46 & 0.45\\
Heliophila digitata & a & 30 & 0.33 & 0.38 & 0.44 & 0.38\\
Heliophila dregeana & p & 17 & 0.33 & 0.37 & 0.33 & 0.32\\
Heliophila elongata & p & 82 & 0.26 & 0.32 & 0.30 & 0.25\\
Heliophila ephemera & a & 3 & 0.14 & 0.27 & 0.31 & 0.26\\
Heliophila esterhuyseniae & p & 3 & 0.21 & 0.30 & 0.37 & 0.27\\
Heliophila eximia & p & 12 & 0.42 & 0.41 & 0.32 & 0.34\\
Heliophila gariepina & a & 12 & 0.50 & 0.53 & 0.48 & 0.41\\
Heliophila glauca & p & 35 & 0.29 & 0.35 & 0.34 & 0.33\\
Heliophila juncea & p & 150 & 0.32 & 0.37 & 0.39 & 0.35\\
Heliophila linearis & p & 94 & 0.32 & 0.33 & 0.28 & 0.30\\
Heliophila macowaniana & a & 31 & 0.33 & 0.38 & 0.44 & 0.39\\
Heliophila macra & p & 22 & 0.30 & 0.30 & 0.32 & 0.29\\
Heliophila macrosperma & p & 5 & 0.28 & 0.36 & 0.35 & 0.25\\
Heliophila minima & p & 35 & 0.36 & 0.45 & 0.51 & 0.39\\
Heliophila namaquana & a & 16 & 0.39 & 0.46 & 0.48 & 0.39\\
Heliophila nubigena & p & 19 & 0.31 & 0.36 & 0.43 & 0.38\\
Heliophila pectinata & a & 16 & 0.27 & 0.34 & 0.50 & 0.34\\
Heliophila polygaloides & p & 12 & 0.40 & 0.48 & 0.42 & 0.34\\
Heliophila pubescens & a & 9 & 0.31 & 0.40 & 0.48 & 0.39\\
Heliophila pusilla & a & 45 & 0.32 & 0.38 & 0.38 & 0.34\\
Heliophila rigidiuscula & p & 201 & 0.30 & 0.33 & 0.28 & 0.24\\
Heliophila scoparia & p & 106 & 0.31 & 0.37 & 0.36 & 0.31\\
Heliophila seselifolia & a & 80 & 0.36 & 0.42 & 0.45 & 0.40\\
Heliophila suavissima & p & 92 & 0.30 & 0.39 & 0.42 & 0.31\\
Heliophila subulata & p & 103 & 0.29 & 0.33 & 0.31 & 0.29\\
Heliophila tricuspidata & p & 8 & 0.28 & 0.33 & 0.38 & 0.30\\
Heliophila trifurca & a & 77 & 0.45 & 0.48 & 0.48 & 0.43\\
Heliophila tulbaghensis & p & 3 & 0.36 & 0.41 & 0.36 & 0.35\\
Heliophila variabilis & a & 97 & 0.35 & 0.41 & 0.40 & 0.37\\
\bottomrule
\addlinespace
\insertTableNotes
\end{longtable}
}
\end{ThreePartTable}
\end{center}

\begin{figure}
\centering
\includegraphics{../figures/maps_drought_examples.pdf}
\caption{\label{fig:mapsdroughtexamples}Example years experiencing contrasting degrees
of drought in southern Africa. Vegetative Health Index (VHI) values
below 40 indicate remotely sensed drought. Drought detection during this
years is validated by precipitation based estimates of drought ocurrance
(Standardized Precipitation Index) which confirm that while (a) 2015 was
one of the worst drought years on record, (b) 1999 was one of the
wettest, and (c) 2012 was typical in terms of precipitation patterns
(Monyela, 2017). It is worth noting that drought can be detected using
the VHI across ecosystems, including those inhabited by perennial rather
than annual \emph{Heliophila} species.}
\end{figure}

\begin{figure}
\centering
\includegraphics{../figures/speciesmaps.pdf}
\caption{\label{fig:speciesmaps}Maps of occurrence records for individual species.
Orange points indicate annual species. Blue points indicate perennial
species.}
\end{figure}

\end{document}